---
title: "Acre EMSI Raster Calculation"
author: "Tyson Lee Swetnam"
date: "2021-03-03"
output: html_document
---

# Setup

## Linux Dependencies

If you are running this for the first time, you may need to install additional geospatial dependencies

```{bash message=FALSE, warning=FALSE}
# sudo add-apt-repository ppa:ubuntugis
# sudo apt-get update
# sudo apt-get install libgdal-dev libgeos-dev libproj-dev
# sudo apt-get install libudunits2-dev
```

## Install missing R Libraries

The following code block executes when the notebook is opened. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("dplyr")) {
   install.packages("dplyr")
   library(dplyr)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}  
if (!require("grid")) {
   install.packages("grid")
   library(grid)
} 
if (!require("lattice")) {
   install.packages("lattice")
   library(lattice)
}
if (!require("leaflet")) {
   install.packages("leaflet")
   library(leaflet)
}
if (!require("lubridate")) {
   install.packages("lubridate")
   library(lubridate)
}
if (!require("magrittr")) {
   install.packages("magrittr")
   library(magrittr)
}
if (!require("raster")) {
   install.packages("raster")
   library(raster)  
}
if (!require("rasterVis")) {
   install.packages("rasterVis")
   library(rasterVis)   
}
if (!require("RColorBrewer")) {
   install.packages("RColorBrewer")
   library(RColorBrewer)
}
if (!require("reshape")) {
   install.packages("reshape")
   library(reshape)
}
if (!require("rgdal")) {
   install.packages("rgdal")
   library(rgdal)   
}  
if (!require("remotes")) {
   install.packages("remotes")
   library(remotes)   
}  
if (!require("PerformanceAnalytics")) {
   install.packages("PerformanceAnalytics")
   library(PerformanceAnalytics)
}
if (!require("scales")) {
   install.packages("scales")
   library(scales)
}
if (!require("sf")) {
   install.packages("sf")
   library(sf)   
}
if (!require("stringr")) {
   install.packages("stringr")
   library(stringr)
}
if (!require("viridis")) {
   install.packages("viridis")
   library(viridis)
}
```

# Transfer datasets from CyVerse

This command uses `wget` to download data into the container.

The CyVerse iRODS data store also supports [WebDav](https://data.cyverse.org); we are hosting the extracted raster data in a folder with anonymous public read-only settings.

All of the landsat data are ~22.8 GB in size, and will take ~18 minutes to transfer within the CyVerse workbench for all four sites.

A subset of the collection will take less time, here we're taking the `/acre` collection. 

```{bash message=FALSE, warning=FALSE}
# uncomment to run
# cd ~/emsi/data/
# time wget -r -nH --cut-dirs=7 --no-parent -l8 --reject="index.html*" https://data.cyverse.org/dav-anon/iplant/home/tswetnam/emsi/data/collections/acre
```

# Import Raster from data directory and check the header
```{r message=FALSE, warning=FALSE}
# Test for raster metadata with GDALinfo
GDALinfo("~/emsi/data/collections/acre/gee-collection-acre-landsat8/LC08_004066_20130611.tif")
```

## Import Raster time series for Landsats 5,7,8
```{r message=FALSE, warning=FALSE}
# Load all rasters in acre-landsat5
setwd("~/emsi/data/collections/acre/gee-collection-acre-landsat5/")
rlist5=list.files(getwd(), pattern=".tif$", full.names=FALSE)
for(i in rlist5) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) } 

# Load all rasters in acre-landsat7
setwd("~/emsi/data/collections/acre/gee-collection-acre-landsat7/")
rlist7=list.files(getwd(), pattern="tif$", full.names=FALSE)
for(i in rlist7) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) }

# Load all rasters in acre-landsat8
setwd("~/emsi/data/collections/acre/gee-collection-acre-landsat8/")
rlist8=list.files(getwd(), pattern="tif$", full.names=FALSE)
for(i in rlist8) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) }
```

```{r}
list_l5 <- ls(pattern="LT05", all.names = TRUE) 
dates_l5 = as.Date(str_sub(list_l5, -8 ,-1), format="%Y%m%d")

list_l7 <- ls(pattern="LE07", all.names = TRUE) 
dates_l7 = as.Date(str_sub(list_l7, -8 ,-1), format="%Y%m%d")

list_l8 <- ls(pattern="LC08", all.names = TRUE) 
dates_l8 = as.Date(str_sub(list_l8, -8 ,-1), format="%Y%m%d")

list_08 <- ls(pattern = "08", all.names = TRUE)
list_08

```

# August (peak greenness) EMSI calc prep
```{r}

lall_08 <- brick(LT05_005066_19980828, 
                    LT05_005066_19990831, 
                    LT05_005066_20000817,
                    LT05_005066_20010820, 
                    LT05_005066_20030826, 
                    LT05_005066_20040828, 
                    LT05_005066_20050831, 
                    LT05_005066_20060802,
                    LT05_005066_20070821, 
                    LT05_005066_20080823, 
                    LT05_005066_20100813,
                    LT05_005066_20110816)


#                    LC08_005066_20130821,
#                    LC08_005066_20140824,
#                    LC08_005066_20150827,
#                    LC08_005066_20160829,
#                    LC08_005066_20170816,
#                    LC08_005066_20180819)
                  

# Calculate mean
#l5_08_mean <- calc(l5_08, mean, na.rm=T)
#l7_08_mean <- calc(l7_08, mean, na.rm=T)
#l8_08_mean <- calc(l8_08, mean, na.rm=T)
lall_08_mean <- calc(lall_08, mean, na.rm=T)

# Calculate sd
#l5_08_sd <- calc(l5_08, sd, na.rm=T)
#l7_08_sd <- calc(l7_08, sd, na.rm=T)
#l8_08_sd <- calc(l8_08, sd, na.rm=T)
lall_08_sd <- calc(lall_08, sd, na.rm=T)


#l5_08_emsi <- overlay(l5_08, l5_08_mean, l5_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

#l7_08_emsi <- overlay(l7_08, l7_08_mean, l7_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

#l8_08_emsi <- overlay(l8_08, l5_08_mean, l5_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

lall_08_emsi <- overlay(lall_08, lall_08_mean, lall_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```

```{r}
#l5_09 <- brick()

# Landsat 8 September dates
#l8_09 <- brick(LC08_035038_20130924,
#                       LC08_035038_20140911,
#                       LC08_035038_20150930,
#                       LC08_035038_20160916,
#                       LC08_035038_20180906)
# Calculate mean
#l8_09_mean <- calc(l8_09, mean)
# Calculate sd
#l8_09_sd <- calc(l8_09, sd)
#l8_09_emsi <- overlay(l8_09, l8_09_mean, l8_09_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```


# Create Leaflet Map of study area 
https://rstudio.github.io/leaflet
http://leafletjs.com/
https://www.r-bloggers.com/interactive-mapping-with-leaflet-in-r/
https://www.color-hex.com/color-palette/19447 

We are going to use a topo map, overlayed with a street map to show states.
To browse all the provider layers,
see http://leaflet-extras.github.io/leaflet-providers/preview/index.html

```{r message=FALSE, warning=FALSE}
# Create custom NDVI color pallete
pal1 <- colorNumeric(c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), values(lall_08_mean), na.color = "transparent")

pal <- colorNumeric(c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), values(lall_08_emsi[[1]]), na.color = "transparent")
```

```{r}
# Scene 127016
m <- leaflet() %>%
  addTiles() %>%
  #addLegend(pal = pal, values = values(lall_08_emsi[[1]]), title = "EMSI") %>%
  #addLegend(pal = pal1, values = values(lall_08_mean), title = "NDVI") %>%
  addRasterImage(lall_08_mean, group = "August Mean NDVI", colors = pal1, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_08_sd, group = "August Standard Deviation NDVI", colors = pal1, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_08_emsi[[1]], group = "August 1998 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_08_emsi[[2]], group = "August 1999 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_08_emsi[[3]], group = "August 2000 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_08_emsi[[4]], group = "August 2001 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_08_emsi[[5]], group = "August 2003 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_08_emsi[[6]], group = "August 2004 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%  
  addRasterImage(lall_08_emsi[[7]], group = "August 2005 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_08_emsi[[8]], group = "August 2006 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%  
  addRasterImage(lall_08_emsi[[9]], group = "August 2007 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_08_emsi[[10]], group = "August 2008 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_08_emsi[[11]], group = "August 2010 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%  
  addRasterImage(lall_08_emsi[[12]], group = "August 2011 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%  
setView(lng = -72.050, lat = -9.128, zoom = 12) %>%
addProviderTiles("Stamen.Toner", group = "Stamen") %>%
addProviderTiles("Esri.WorldImagery", group = "ESRI Satellite", options = providerTileOptions(opacity = 0.66, transparent = TRUE)) %>%
addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
#layers control panel
addLayersControl(baseGroups = c("Stamen", "ESRI Satellite", "OpenStreetMap"), overlayGroups = c("August Mean NDVI", "August Standard Deviation NDVI", "August 1998 EMSI", "August 1999 EMSI", "August 2000 EMSI", "August 2001 EMSI", "August 2003 EMSI", "August 2004 EMSI", "August 2005 EMSI", "August 2006 EMSI", "August 2007 EMSI", "August 2008 EMSI", "August 2010 EMSI", "August 2011 EMSI"), options = layersControlOptions(collapsed = TRUE))

# "August 2013 EMSI", "August 2014 EMSI", "August 2015 EMSI", "August 2016 EMSI", "August 2017 EMSI", "August 2018 EMSI"
m
```


```{r}
year_id <- c('LT05_005066_19980828' = "1998", 
                    'LT05_005066_19990831' = "1999", 
                    'LT05_005066_20000817' = "2000",
                    'LT05_005066_20010820' = "2001", 
                    'LT05_005066_20030826' = "2003", 
                    'LT05_005066_20040828' = "2004", 
                    'LT05_005066_20050831' = "2005", 
                    'LT05_005066_20060802' = "2006",
                    'LT05_005066_20070821' = "2007", 
                    'LT05_005066_20080823' = "2008", 
                    'LT05_005066_20100813' = "2010",
                    'LT05_005066_20110816' = "2011")


year_ids <- c('layer.1'="1998",
              'layer.2'="1999",
              'layer.3'="2000",
              'layer.4'="2001",
              'layer.5'="2003",
              'layer.6'="2004",
              'layer.7'="2005",
              'layer.8'="2006",
              'layer.9'="2007",
              'layer.10'="2008",
              'layer.11'="2010",
              'layer.12'="2011")
```

```{r}
## Multipanel graph Summer 2000 - 2013
lall_stack <- stack(LT05_005066_19980828, 
                    LT05_005066_19990831, 
                    LT05_005066_20000817,
                    LT05_005066_20010820, 
                    LT05_005066_20030826, 
                    LT05_005066_20040828, 
                    LT05_005066_20050831, 
                    LT05_005066_20060802,
                    LT05_005066_20070821, 
                    LT05_005066_20080823, 
                    LT05_005066_20100813,
                    LT05_005066_20110816)

lall_stack_df <- as.data.frame(lall_stack, xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = lall_stack_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  facet_wrap(~ variable, labeller = as_labeller(year_id), ncol = 6) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
 
ggsave('~/emsi/rmd/figures/ndvi_timeseries_acre.png', width = 12, dpi = 300, bg = "transparent")
```

```{r}
## Multipanel graph Augusts 2000 - 2018
lall_stack <- stack(LT05_005066_19980828, 
                    LT05_005066_19990831, 
                    LT05_005066_20000817,
                    LT05_005066_20010820, 
                    LT05_005066_20030826, 
                    LT05_005066_20040828, 
                    LT05_005066_20050831, 
                    LT05_005066_20060802,
                    LT05_005066_20070821, 
                    LT05_005066_20080823, 
                    LT05_005066_20100813,
                    LT05_005066_20110816)

lall_stack_df <- as.data.frame(lall_stack, xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = lall_stack_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  facet_wrap(~ variable, labeller = as_labeller(year_id), ncol = 6) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
 
ggsave('~/emsi/rmd/figures/ndvi_timeseries_acre.png', width = 22, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2010_stack_ndvi_df <- as.data.frame(lall_stack[[11]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2010_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/emsi/rmd/figures/ndvi_2010_acre.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2006_stack_ndvi_df <- as.data.frame(lall_stack[[8]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2006_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/emsi/rmd/figures/ndvi_2006_acre.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```



```{r}
# Calculate mean
lall_stack_mean <- calc(lall_stack, mean, na.rm=T)
# Calculate sd
lall_stack_sd <- calc(lall_stack, sd, na.rm=T)
lall_stack_emsi <- overlay(lall_stack, lall_stack_mean, lall_stack_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```


```{r}
lall_stack_emsi_df <- as.data.frame(lall_stack_emsi, xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = lall_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-2.5,2.7), guide = guide_colorbar(title ="EMSI")) +
  facet_wrap(~ variable, labeller = as_labeller(year_ids), ncol = 6) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/emsi/rmd/figures/emsi_timeseries_acre.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2010_stack_emsi_df <- as.data.frame(lall_stack_emsi[[11]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2010_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3,2.7), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/emsi/rmd/figures/emsi_2010_acre.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2006_stack_emsi_df <- as.data.frame(lall_stack_emsi[[8]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2006_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3.2,2.7), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/emsi/rmd/figures/emsi_2006_acre.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2016_stack_emsi_df <- as.data.frame(lall_stack_emsi[[24]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2016_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3,3.2), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/emsi/rmd/figures/emsi_2016_acre.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```


```{r}
ggplot(lall_stack_emsi_df) + geom_histogram(aes(value, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-2.5,2.7), guide = guide_colorbar(title ="EMSI")) +
  ylab("Density") + xlab("EMSI") + ggtitle("August") +
  facet_wrap(~variable, labeller = as_labeller(year_ids), ncol = 6) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
```