---
title: "Chiricahuas ERC vs EMSI"
author: "Tyson Lee Swetnam <tswetnam@arizona.edu>"
date: "2020-08-17"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
always_allow_html: yes
---

# Setup

## Linux Dependencies

If you are running this on a VM outside CyVerse, you may need to install additional dependencies

```{bash message=FALSE, warning=FALSE}
# sudo add-apt-repository ppa:ubuntugis
# sudo apt-get update
# sudo apt-get install libgdal-dev libgeos-dev libproj-dev
# sudo apt-get install libudunits2-dev
```

## Using iCommands

If you haven't used iRODS iCommands before, you'll need to set up a CyVerse account. 

In the running RStudio container open a terminal window and initiate a connection to the CyVerse Data Store.

Below is the example from Terminal:

```
$ iinit
One or more fields in your iRODS environment file (irods_environment.json) are
missing; please enter them.
Enter the host name (DNS) of the server to connect to: data.cyverse.org
Enter the port number: 1247
Enter your irods user name: tswetnam
Enter your irods zone: iplant
Those values will be added to your environment file (for use by
other iCommands) if the login succeeds.

Enter your current iRODS password:
```

This will generate an iRODS environment JSON file located in a hidden folder in your user home directory

```
/home/rstudio/.irods/irods_environment.json
```

The contents of which are:

```
{
    "irods_host": "data.cyverse.org",
    "irods_port": 1247,
    "irods_user_name": "tswetnam",
    "irods_zone_name": "iplant"
}
```

Once this `irods_environment.json` exists, you'll be able to view your data store contents by typing the `ils` command.


## Install missing R Libraries

The following code block executes when the notebook is opened. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("dplyr")) {
   install.packages("dplyr")
   library(dplyr)
}
if (!require("gridExtra")) {
   install.packages("gridExtra")
   library(gridExtra)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}  
if (!require("grid")) {
   install.packages("grid")
   library(grid)
} 
if (!require("lattice")) {
   install.packages("lattice")
   library(lattice)
}
if (!require("leaflet")) {
   install.packages("leaflet")
   library(leaflet)
}
if (!require("lubridate")) {
   install.packages("lubridate")
   library(lubridate)
}
if (!require("magrittr")) {
   install.packages("magrittr")
   library(magrittr)
}
if (!require("raster")) {
   install.packages("raster")
   library(raster)  
}
if (!require("rasterVis")) {
   install.packages("rasterVis")
   library(rasterVis)   
}
if (!require("RColorBrewer")) {
   install.packages("RColorBrewer")
   library(RColorBrewer)
}
if (!require("reshape")) {
   install.packages("reshape")
   library(reshape)
}
if (!require("rgdal")) {
   install.packages("rgdal")
   library(rgdal)   
}  
if (!require("remotes")) {
   install.packages("remotes")
   library(remotes)   
}  
if (!require("PerformanceAnalytics")) {
   install.packages("PerformanceAnalytics")
   library(PerformanceAnalytics)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("scales")) {
   install.packages("scales")
   library(scales)
}
if (!require("sf")) {
   install.packages("sf")
   library(sf)   
}
if (!require("stringr")) {
   install.packages("stringr")
   library(stringr)
}
if (!require("viridis")) {
   install.packages("viridis")
   library(viridis)
}
```

# Transfer datasets from CyVerse

This command uses `wget` to download data into the container.

The CyVerse iRODS data store also supports [WebDav](https://data.cyverse.org); we are hosting the extracted raster data in a folder with anonymous public read-only settings.

Data are ~22.8 GB in size, and will take ~18 minutes to transfer within the CyVerse workbench.

```{bash message=FALSE, warning=FALSE}
# uncomment to run
# cd ~/emsi/data/
# time wget -r -nH --cut-dirs=6 --no-parent -l8 --reject="index.html*" https://data.cyverse.org/dav-anon/iplant/home/tswetnam/emsi/data/collections/
```

## Data

I extracted the entire time series of 16-day Landsat  NDVI data from Google Earth Engine (GEE) for an AOI in the Chiricahua Mountains, Arizona. 

I also extracted the Gridmet daily product for fire behavior.

In GEE I filtered the time series using the Landsat Surface Reflectance Tier 1 data.

```{r message=FALSE, warning=FALSE}

## Import Gridmet 1000-hour Fuel Moistures, Energy Release Component (ERC), and Burning Index (BI)

## 1000hr Fuel Moistures
fm1000_1984_1994 <- read.csv("~/emsi/data/gridmet_fm1000_chiricahua_sa1_1984_1994.csv")
fm1000_1995_2005 <- read.csv("~/emsi/data/gridmet_fm1000_chiricahua_sa1_1995_2005.csv")
fm1000_2006_2018 <- read.csv("~/emsi/data/gridmet_fm1000_chiricahua_sa1_2006_2018.csv")

fm1000 <- rbind(fm1000_1984_1994, fm1000_1995_2005, fm1000_2006_2018)
## Convert the dates to an R readable format
fm1000$asdate <- as.Date(fm1000$date, format = "%b %d, %Y")
fm1000$julian <- yday(fm1000$asdate)
fm1000$julian_rounded <- round((fm1000$julian/365)*52)*7

## ERC
erc_1984_1994 <- read.csv("~/emsi/data/gridmet_erc_chiricahua_sa1_1984_1994.csv")
erc_1995_2005 <- read.csv("~/emsi/data/gridmet_erc_chiricahua_sa1_1995_2005.csv")
erc_2006_2018 <- read.csv("~/emsi/data/gridmet_erc_chiricahua_sa1_2006_2018.csv")

erc <- rbind(erc_1984_1994, erc_1995_2005, erc_2006_2018)
## Convert the dates to an R readable format
erc$asdate <- as.Date(erc$date, format = "%b %d, %Y")
erc$julian <- yday(erc$asdate)
erc$julian_rounded <- round((erc$julian/365)*52)*7

## BI
## 1000hr Fuel Moistures
bi_1984_1994 <- read.csv("~/emsi/data/gridmet_bi_chiricahua_sa1_1984_1994.csv")
bi_1995_2005 <- read.csv("~/emsi/data/gridmet_bi_chiricahua_sa1_1995_2005.csv")
bi_2006_2018 <- read.csv("~/emsi/data/gridmet_bi_chiricahua_sa1_2006_2018.csv")

bi <- rbind(bi_1984_1994, bi_1995_2005, bi_2006_2018)
## Convert the dates to an R readable format
bi$asdate <- as.Date(bi$date, format = "%b %d, %Y")
bi$julian <- yday(bi$asdate)
bi$julian_rounded <- round((bi$julian/365)*52)*7

## Import the Landsat 16-day NDVI from Google Earth Engine extract
l8_sa2 <- read.csv("~/emsi/data/landsat8_chiricahua_sa2_ndvi.csv")
l7_sa2 <- read.csv("~/emsi/data/landsat7_chiricahua_sa2_ndvi.csv")
l5_sa2 <- read.csv("~/emsi/data/landsat5_chiricahua_sa2_ndvi.csv")
l8_sa1 <- read.csv("~/emsi/data/landsat8_chiricahua_sa1_ndvi.csv")
l7_sa1 <- read.csv("~/emsi/data/landsat7_chiricahua_sa1_ndvi.csv")
l5_sa1 <- read.csv("~/emsi/data/landsat5_chiricahua_sa1_ndvi.csv")

## Remove days with low NDVI values - likely due to cluds and snow.
l8_sa2$ndvi_nans <- as.numeric(as.character(l8_sa2$ndvi))
l8_sa2$ndvi_range <- ifelse(l8_sa2$ndvi_nans>0.15,l8_sa2$ndvi_nans,NA)
## Convert the dates to an R readable format
l8_sa2$asdate <- as.Date(l8_sa2$date, format = "%b %d, %Y")
l8_sa2$julian <- yday(l8_sa2$asdate)
l8_sa2$julian_rounded <- round((l8_sa2$julian/365)*52)*7

l7_sa2$ndvi_nans <- as.numeric(as.character(l7_sa2$ndvi))
l7_sa2$ndvi_range <- ifelse(l7_sa2$ndvi_nans>0.15,l7_sa2$ndvi_nans,NA)
## Convert the dates to an R readable format
l7_sa2$asdate <- as.Date(l7_sa2$date, format = "%b %d, %Y")
l7_sa2$julian <- yday(l7_sa2$asdate)
l7_sa2$julian_rounded <- round((l7_sa2$julian/365)*52)*7


l5_sa2$ndvi_nans <- as.numeric(as.character(l5_sa2$ndvi))
l5_sa2$ndvi_range <- ifelse(l5_sa2$ndvi_nans>0.3,l5_sa2$ndvi_nans,NA)
## Convert the dates to an R readable format
l5_sa2$asdate <- as.Date(l5_sa2$date, format = "%b %d, %Y")
l5_sa2$julian <- yday(l5_sa2$asdate)
l5_sa2$julian_rounded <- round((l5_sa2$julian/365)*52)*7

l8_sa1$ndvi_nans <- as.numeric(as.character(l8_sa1$ndvi))
l8_sa1$ndvi_range <- ifelse(l8_sa1$ndvi_nans>0.3,l8_sa1$ndvi_nans,NA)
## Convert the dates to an R readable format
l8_sa1$asdate <- as.Date(l8_sa1$date, format = "%b %d, %Y")
l8_sa1$julian <- yday(l8_sa1$asdate)
l8_sa1$julian_rounded <- round((l8_sa1$julian/365)*52)*7

l7_sa1$ndvi_nans <- as.numeric(as.character(l7_sa1$ndvi))
l7_sa1$ndvi_range <- ifelse(l7_sa1$ndvi_nans>0.3,l7_sa1$ndvi_nans,NA)
## Convert the dates to an R readable format
l7_sa1$asdate <- as.Date(l7_sa1$date, format = "%b %d, %Y")
l7_sa1$julian <- yday(l7_sa1$asdate)
l7_sa1$julian_rounded <- round((l7_sa1$julian/365)*52)*7

l5_sa1$ndvi_nans <- as.numeric(as.character(l5_sa1$ndvi))
l5_sa1$ndvi_range <- ifelse(l5_sa1$ndvi_nans>0.35,l5_sa1$ndvi_nans,NA)
## Convert the dates to an R readable format
l5_sa1$asdate <- as.Date(l5_sa1$date, format = "%b %d, %Y")
l5_sa1$julian <- yday(l5_sa1$asdate)
l5_sa1$julian_rounded <- round((l5_sa1$julian/365)*52)*7

```

Plot NDVI:
```{r message=FALSE, warning=FALSE}
# Plot examples for a range of dates

# Landsat 5 Study Area 1
range_dates = l5_sa1[1:478,]
p1 <- ggplot(range_dates[!is.na(range_dates$ndvi_range),], aes(x=asdate, y=ndvi_range))+ geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 5 1984-2011 16-Day NDVI Chiricahua Study Area 1") + xlab("Date") + ylab("NDVI") + geom_line(color='blue', size = 0.1) 

p1 <- p1 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggplotly(p1)
plot(p1)
ggsave('~/emsi/rmd/figures/landsat5_chiricahua_sa1_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")
```

```{r}
# Landsat 5 Study Area 2
range_dates = l5_sa2[1:478,]
p2 <- ggplot(range_dates[!is.na(range_dates$ndvi_range),], aes(x=asdate, y=ndvi_range))+ geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 5 1984-2011 16-Day NDVI Chiricahua Study Area 2") + xlab("Date") + ylab("NDVI") + geom_line(color='blue', size = 0.1) 

p2 <- p2 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggplotly(p2)
plot(p2)
ggsave('~/emsi/rmd/figures/landsat5_chiricahua_sa2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

```

```{r message=FALSE, warning=FALSE}
# Plot examples for a range of dates

# Landsat 7 Study Area 1
range_dates = l7_sa1[1:403,]
p1 <- ggplot(range_dates[!is.na(range_dates$ndvi_range),], aes(x=asdate, y=ndvi_range))+ geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 7 1999-2018 16-Day NDVI Chiricahua Study Area 1") + xlab("Date") + ylab("NDVI") + geom_line(color='blue', size = 0.1) 

p1 <- p1 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggplotly(p1)
plot(p1)
ggsave('~/emsi/rmd/figures/landsat7_chiricahua_sa1_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

# Landsat 7 Study Area 2
range_dates = l7_sa2[1:403,]
p2 <- ggplot(range_dates[!is.na(range_dates$ndvi_range),], aes(x=asdate, y=ndvi_range))+ geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 7 1999-2018 16-Day NDVI Chiricahua Study Area 2") + xlab("Date") + ylab("NDVI") + geom_line(color='blue', size = 0.1) 

p2 <- p2 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggplotly(p2)
plot(p2)
ggsave('~/emsi/rmd/figures/landsat7_chiricahua_sa2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

```

```{r message=FALSE, warning=FALSE}
# Plot examples for a range of dates

# Landsat 8 Study Area 1
range_dates = l8_sa1[1:111,]
p1 <- ggplot(range_dates[!is.na(range_dates$ndvi_range),], aes(x=asdate, y=ndvi_range))+ geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 8 2013-2018 16-Day NDVI Chiricahua Study Area 1") + xlab("Date") + ylab("NDVI") + geom_line(color='blue', size = 0.1) 

p1 <- p1 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggplotly(p1)
plot(p1)
ggsave('~/emsi/rmd/figures/landsat8_chiricahua_sa1_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")
```

```{r}
# Landsat 7 Study Area 2
range_dates = l8_sa2[1:111,]
p2 <- ggplot(range_dates[!is.na(range_dates$ndvi_range),], aes(x=asdate, y=ndvi_range))+ geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 8 2013-2018 16-Day NDVI Chiricahua Study Area 2") + xlab("Date") + ylab("NDVI") + geom_line(color='blue', size = 0.1) 

p2 <- p2 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggplotly(p2)
plot(p2)
ggsave('~/emsi/rmd/figures/landsat7_chiricahua_sa2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

```

In order to make EMSI sensitive to inter-annual variation, e.g. vegetation can be more stressed than average during fall or spring dry seasons, I calculate a moving average over the Julian calendar year (1 - 365(6) days) for the time series I had available from Landsat 8.

I used Loess function (curved blue line) to find the moving average of each period. The black line links the means of each 16-day period. 

```{r message=FALSE, warning=FALSE}
# Landsat 8 Chiricahua Mean NDVI

p2 <- ggplot(l8_sa1[1:111,], aes(x=julian_rounded,y=ndvi_range)) + 
  geom_point(aes(color=ndvi_range)) + 
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + 
  ggtitle("Landsat 8 Reference 16-Day Moving Average NDVI over Chiricahua") + 
  xlab("Julian Date") + 
  ylab("NDVI") + 
  geom_smooth(method="loess", se=TRUE, level=0.95) 

p2 <- p2 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p2)
ggsave('~/emsi/rmd/figures/landsat8_chiricahua_ndvi_mean.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(p2)

```

Here I calculate EMSI during a reference period between 1984 and 1996

```{r message=FALSE, warning=FALSE}

# Predict the Julian date average of Landsat 8 NDVI using a Loess function
mean_loess5 <- predict(loess(ndvi_range ~ julian_rounded, l8_sa1[1:111,], span=0.1), l8_sa1[1:111,]$julian_rounded)

# Calculate mean NDVI for each 16-day Julian period
ndvi_mean5 <- setNames(aggregate(l8_sa1[1:111,]$ndvi_range, list(l8_sa1[1:111,]$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))

# Calculate emsi using Performance Analytics sd.multiperiod
l8_sa2$emsi = (l8_sa2$ndvi_range - mean_loess5) /  sd.multiperiod(l8_sa2$ndvi_range,scale=1)
l7_sa2$emsi = (l7_sa2$ndvi_range - mean_loess5) /  sd.multiperiod(l7_sa2$ndvi_range,scale=1)
l5_sa2$emsi = (l5_sa2$ndvi_range - mean_loess5) /  sd.multiperiod(l5_sa2$ndvi_range,scale=1)
l8_sa1$emsi = (l8_sa1$ndvi_range - mean_loess5) /  sd.multiperiod(l8_sa1$ndvi_range,scale=1)
l7_sa1$emsi = (l7_sa1$ndvi_range - mean_loess5) /  sd.multiperiod(l7_sa1$ndvi_range,scale=1)
l5_sa1$emsi = (l5_sa1$ndvi_range - mean_loess5) /  sd.multiperiod(l5_sa1$ndvi_range,scale=1)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Histograms of the time series
hist(l8_sa1[1:100,]$julian_rounded,breaks=52,xlab="Julian day of year", main = "Landsat 8 Histogram of Dates in Reference Dataset")
```

The calculation of emsi still retains some of the same trends as the original NDVI, but it is now sensitive to changes in vegetation condition during the winter, spring, and fall seasons.

```{r echo=FALSE, message=FALSE, warning=FALSE}
range_dates = l7_sa1[1:403,]

p4 <- ggplot(range_dates[!is.na(range_dates$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-5.0,5.0), breaks = seq(-5,5,0.5)) + 
  ggtitle("Landsat 7 EMSI") + 
  xlab("Date") + 
  ylab("EMSI") + 
  geom_line(size = 0.1, color="blue")

p4 <- p4 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p4)
ggsave('~/emsi/rmd/figures/landsat7_chiricahua_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(p4)
```

I also show the 16-day average of NDVI over the 18 year time series (early 1984 until 2017) for all three satellites.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(l7_sa1[!is.na(l7_sa1$emsi),], aes(x=asdate,y=emsi)) + geom_point(aes(color=emsi)) + scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + scale_y_continuous(limits = c(-5,5), breaks = seq(-5,5,0.5)) + ggtitle("Landsat 7 High Peak EMSI") + xlab("Date") + ylab("EMSI") + geom_line(size = 0.1, color="blue")

ggplot(l5_sa1[!is.na(l5_sa1$emsi),], aes(x=asdate,y=emsi)) + geom_point(aes(color=emsi)) + scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + scale_y_continuous(limits = c(-5,5), breaks = seq(-5,5,0.5)) + ggtitle("Landsat 5 High Peak EMSI") + xlab("Date") + ylab("EMSI") + geom_line(size = 0.1)

```

Alternately, I can plot the three time series together.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot time series
l8 <- l8_sa1[1:111,]
l7 <- l7_sa1[1:403,]
l5 <- l5_sa1[1:486,]

p2 <- ggplot(l5, aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi), size = 1) + 
  scale_colour_gradient2(low = "red", high = "green", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-5,5), breaks = seq(-5,5,0.5)) +
  geom_line(data = l8, aes(x=asdate,y=emsi), size = 0.1, color = "black") + 
  geom_line(data = l5, aes(x=asdate,y=emsi), size = 0.1, color = "black") + 
  geom_line(data = l7, aes(x=asdate,y=emsi), size = 0.1, color = "blue") +
  ggtitle("Chiricahua Study Area 1 EMSI") + 
  xlab("Calendar Date") + ylab("EMSI")


p2 <- p2 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggsave('~/emsi/rmd/figures/landsat_chiricahua_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

plot(p2)
ggplotly(p2)
```
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot time series
l8 <- l8_sa2[1:111,]
l7 <- l7_sa2[1:403,]
l5 <- l5_sa2[1:486,]

p2 <- ggplot(l5, aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi), size = 1) + 
  scale_colour_gradient2(low = "red", high = "green", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-5,5), breaks = seq(-5,5,0.5)) +
  geom_line(data = l8, aes(x=asdate,y=emsi), size = 0.1, color = "black") + 
  geom_line(data = l5, aes(x=asdate,y=emsi), size = 0.1, color = "black") + 
  geom_line(data = l7, aes(x=asdate,y=emsi), size = 0.1, color = "blue") +
  ggtitle("Chiricahua Study Area 1 EMSI") + 
  xlab("Calendar Date") + ylab("EMSI")


p2 <- p2 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggsave('~/emsi/rmd/figures/landsat_chiricahua_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

plot(p2)
ggplotly(p2)
```

Plotting the GridMet data vs the EMSI

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Loess function of FM1000
p <- ggplot(fm1000[9131:10591,], aes(x=asdate,y=fm1000)) + 
  geom_point(aes(color=fm1000), size = 0.1) + 
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 50) + 
  ggtitle("Fuel Moistures 1000 hour Moving Average Chiricahua") + 
  xlab("Date") + 
  ylab("Fuel Moisture") + 
  geom_smooth(method="loess", span = 0.01, size = 0.1)

p <- p + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
) + scale_y_continuous(position = "right")

plot(p)
ggsave('~/emsi/rmd/figures/landsat_chiricahua_fm1000.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

```



```{r echo=FALSE, message=FALSE, warning=FALSE}
# Loess function of ERC

p <- ggplot(erc[9131:10591,], aes(x=asdate,y=erc)) + 
  geom_point(aes(color=erc), size = 0.1) + 
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 50) + 
  ggtitle("ERC Moving Average Chiricahua") + 
  xlab("Date") + 
  ylab("Energy Release Component") + 
  geom_smooth(method="loess", span = 0.01, size = 0.1)

p <- p + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
) + scale_y_continuous(position = "right")

plot(p)
ggsave('~/emsi/rmd/figures/landsat_chiricahua_erc.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Loess function of Burning Indexx

p <- ggplot(bi[9131:10591,], aes(x=asdate,y=bi)) + 
  geom_point(aes(color=bi), size = 0.1) + 
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 50) + 
  ggtitle("Burning Index Moving Average Chiricahua") + 
  xlab("Date") + 
  ylab("Buring Index") + 
  geom_smooth(method="loess", span = 0.01, size = 0.1)

p <- p + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
) + scale_y_continuous(position = "right")

plot(p)
ggsave('~/emsi/rmd/figures/landsat_chiricahua_bi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

```

```{r}
range_dates = l7_sa2[206:293,]

p4 <- ggplot(range_dates[!is.na(range_dates$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-4.0,3.5), breaks = seq(-4,3.5,0.5)) + 
  ggtitle("Landsat 7 EMSI") + 
  xlab("Date") + 
  ylab("EMSI") + 
  geom_line(size = 0.1, color="blue")

p4 <- p4 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p4)

ggsave('~/emsi/rmd/figures/landsat_chiricahua_emsi_l7_sr2.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

```

```{r}
# LANDSAT 5 NDVI
range_dates = l5_sa2[100:161,]

p4 <- ggplot(range_dates[!is.na(range_dates$ndvi),], aes(x=asdate,y=ndvi)) + 
  geom_point(aes(color=ndvi)) + 
  scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.4) + 
  scale_y_continuous(limits = c(0.2,0.6), breaks = seq(0.2,0.6,0.1)) + 
  ggtitle("Landsat 5 NDVI") + 
  xlab("Date") + 
  ylab("NDVI") + 
  geom_line(size = 0.1, color="blue")

p4 <- p4 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_blank(), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p4)

ggsave('~/emsi/rmd/figures/landsat_chiricahua_ndvi_l5_sr2.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
range_dates = l8_sa1[1:111,]

p2 <- ggplot(range_dates[!is.na(range_dates$emsi),], aes(x=asdate, y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0) + 
  ggtitle("Landsat 8 2013-2018 16-Day NDVI Chiricahua Study Area 1") + 
  xlab("Date") + 
  ylab("EMSI") + 
  geom_line(color='blue', size = 0.1) 

p2 <- p2 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggplotly(p2)
```