---
title: "EMSI LANDSAT 5,7 Pinalenos"
author: "Tyson Lee Swetnam <tswetnam@arizona.edu>"
date: "2020-08-17"
output:
  pdf_document: default
  html_document: default
editor_options: 
  chunk_output_type: console
always_allow_html: yes
---

# Setup

## Linux Dependencies

If you are running this on a VM outside CyVerse, you may need to install additional dependencies

```{bash message=FALSE, warning=FALSE}
# sudo add-apt-repository ppa:ubuntugis
# sudo apt-get update
# sudo apt-get install libgdal-dev libgeos-dev libproj-dev
# sudo apt-get install libudunits2-dev
```

## Using iCommands

If you haven't used iRODS iCommands before, you'll need to set up a CyVerse account. 

In the running RStudio container open a terminal window and initiate a connection to the CyVerse Data Store.

Below is the example from Terminal:

```
$ iinit
One or more fields in your iRODS environment file (irods_environment.json) are
missing; please enter them.
Enter the host name (DNS) of the server to connect to: data.cyverse.org
Enter the port number: 1247
Enter your irods user name: tswetnam
Enter your irods zone: iplant
Those values will be added to your environment file (for use by
other iCommands) if the login succeeds.

Enter your current iRODS password:
```

This will generate an iRODS environment JSON file located in a hidden folder in your user home directory

```
/home/rstudio/.irods/irods_environment.json
```

The contents of which are:

```
{
    "irods_host": "data.cyverse.org",
    "irods_port": 1247,
    "irods_user_name": "tswetnam",
    "irods_zone_name": "iplant"
}
```

Once this `irods_environment.json` exists, you'll be able to view your data store contents by typing the `ils` command.


## Install missing R Libraries

The following code block executes when the notebook is opened. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("dplyr")) {
   install.packages("dplyr")
   library(dplyr)
}
if (!require("gridExtra")) {
   install.packages("gridExtra")
   library(gridExtra)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}  
if (!require("grid")) {
   install.packages("grid")
   library(grid)
} 
if (!require("lattice")) {
   install.packages("lattice")
   library(lattice)
}
if (!require("leaflet")) {
   install.packages("leaflet")
   library(leaflet)
}
if (!require("lubridate")) {
   install.packages("lubridate")
   library(lubridate)
}
if (!require("magrittr")) {
   install.packages("magrittr")
   library(magrittr)
}
if (!require("raster")) {
   install.packages("raster")
   library(raster)  
}
if (!require("rasterVis")) {
   install.packages("rasterVis")
   library(rasterVis)   
}
if (!require("RColorBrewer")) {
   install.packages("RColorBrewer")
   library(RColorBrewer)
}
if (!require("reshape")) {
   install.packages("reshape")
   library(reshape)
}
if (!require("rgdal")) {
   install.packages("rgdal")
   library(rgdal)   
}  
if (!require("remotes")) {
   install.packages("remotes")
   library(remotes)   
}  
if (!require("PerformanceAnalytics")) {
   install.packages("PerformanceAnalytics")
   library(PerformanceAnalytics)
}
if (!require("plotly")) {
   install.packages("plotly")
   library(plotly)
}
if (!require("scales")) {
   install.packages("scales")
   library(scales)
}
if (!require("sf")) {
   install.packages("sf")
   library(sf)   
}
if (!require("stringr")) {
   install.packages("stringr")
   library(stringr)
}
if (!require("viridis")) {
   install.packages("viridis")
   library(viridis)
}
```

# Transfer datasets from CyVerse

This command uses `wget` to download data into the container.

The CyVerse iRODS data store also supports [WebDav](https://data.cyverse.org); we are hosting the extracted raster data in a folder with anonymous public read-only settings.

Data are ~22.8 GB in size, and will take ~18 minutes to transfer within the CyVerse workbench.

```{bash message=FALSE, warning=FALSE}
# uncomment to run
# cd ~/emsi/data/
# time wget -r -nH --cut-dirs=6 --no-parent -l8 --reject="index.html*" https://data.cyverse.org/dav-anon/iplant/home/tswetnam/emsi/data/collections/
```

## Data

I extracted the entire time series of 16-day Landsat  NDVI data from Google Earth Engine (GEE) for an AOI on the Cienega Ranch, near the Mt Graham International Observatory (MGIO) and nearby High Peak (Mt. Graham).

In GEE I filtered the time series using the Landsat Surface Reflectance Tier 1 data.

```{r message=FALSE, warning=FALSE}
## Import the Landsat 16-day NDVI from Google Earth Engine extract

aoi4 <- read.csv("/Users/tyson_swetnam/emsi/data/landsat7_mgio_ndvi.csv")
aoi3 <- read.csv("/Users/tyson_swetnam/emsi/data/landsat5_mgio_ndvi.csv")
aoi2 <- read.csv("/Users/tyson_swetnam/emsi/data/landsat7_highpeak_ndvi.csv")
aoi1 <- read.csv("/Users/tyson_swetnam/emsi/data/landsat5_highpeak_ndvi.csv")

## Remove days with low NDVI values - likely due to cluds and snow.

aoi4$ndvi_nans <- as.numeric(as.character(aoi4$ndvi))
aoi4$ndvi_range <- ifelse(aoi4$ndvi_nans>0,aoi4$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi4$asdate <- as.Date(aoi4$date, format = "%b - %d - %Y")
aoi4$julian <- yday(aoi4$asdate)
aoi4$julian_rounded <- round((aoi4$julian/365)*52)*7

aoi3$ndvi_nans <- as.numeric(as.character(aoi3$ndvi))
aoi3$ndvi_range <- ifelse(aoi3$ndvi_nans>0.3,aoi3$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi3$asdate <- as.Date(aoi3$date, format = "%b - %d - %Y")
aoi3$julian <- yday(aoi3$asdate)
aoi3$julian_rounded <- round((aoi3$julian/365)*52)*7

aoi2$ndvi_nans <- as.numeric(as.character(aoi2$ndvi))
aoi2$ndvi_range <- ifelse(aoi2$ndvi_nans>0,aoi2$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi2$asdate <- as.Date(aoi2$date, format = "%b - %d - %Y")
aoi2$julian <- yday(aoi2$asdate)
aoi2$julian_rounded <- round((aoi2$julian/365)*52)*7

aoi1$ndvi_nans <- as.numeric(as.character(aoi1$ndvi))
aoi1$ndvi_range <- ifelse(aoi1$ndvi_nans>0,aoi1$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi1$asdate <- as.Date(aoi1$date, format = "%b - %d - %Y")
aoi1$julian <- yday(aoi1$asdate)
aoi1$julian_rounded <- round((aoi1$julian/365)*52)*7

```

Plot NDVI:
```{r message=FALSE, warning=FALSE}
# Plot examples for a range of dates
range_dates = aoi3[1:478,]
p1 <- ggplot(range_dates[!is.na(range_dates$ndvi_range),], aes(x=asdate, y=ndvi_range))+ geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 5 1984-2011 16-Day NDVI MGIO") + xlab("Date") + ylab("NDVI") + geom_line(color='white', size = 0.1) 

p1 <- p1 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggplotly(p1)
plot(p1)

ggsave('~/Documents/landsat5_mgio_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")
```

In order to make EMSI sensitive to inter-annual variation, e.g. vegetation can be more stressed than average during fall or spring dry seasons, I calculate a moving average over the Julian calendar year (1 - 365(6) days) for the time series I had available from Landsat 7.

Here I calculate EMSI during a reference period between 1984 and 1996

```{r message=FALSE, warning=FALSE}

# Predict the Julian date average of Landsat 7 NDVI using a Loess function
mean_loess7 <- predict(loess(ndvi_range ~ julian_rounded, aoi3[1:196,], span=0.1), aoi3[1:196,]$julian_rounded)
# Calculate mean NDVI for each 16-day Julian period
ndvi_mean7 <- setNames(aggregate(aoi3[1:196,]$ndvi_range, list(aoi3[1:196,]$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))

# Calculate emsi using Performance Analytics sd.multiperiod
aoi4$emsi = (aoi4$ndvi_range - mean_loess7) /  sd.multiperiod(aoi4$ndvi_range,scale=1)
aoi3$emsi = (aoi3$ndvi_range - mean_loess7) /  sd.multiperiod(aoi3$ndvi_range,scale=1)
aoi2$emsi = (aoi2$ndvi_range - mean_loess7) /  sd.multiperiod(aoi2$ndvi_range,scale=1)
aoi1$emsi = (aoi1$ndvi_range - mean_loess7) /  sd.multiperiod(aoi1$ndvi_range,scale=1)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Histograms of the time series
hist(aoi3[1:196,]$julian_rounded,breaks=52,xlab="Julian day of year", main = "Landsat 5 Histogram of Dates in Reference Dataset")
```

I used Loess function (curved blue line) to find the moving average of each period. The black line links the means of each 16-day period. 

```{r message=FALSE, warning=FALSE}
# Landsat 5 MGIO

p2 <- ggplot(aoi3[1:196,], aes(x=julian_rounded,y=ndvi_range)) + 
  geom_point(aes(color=ndvi_range)) + 
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + 
  ggtitle("Landsat 5 1984-1996 Reference 16-Day Moving Average NDVI over MGIO") + 
  xlab("Julian Date") + 
  ylab("NDVI") + 
  geom_smooth(method="loess", se=TRUE, level=0.95) 

p2 <- p2 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p2)
ggsave('~/Documents/landsat5_mgio_ndvi_mean.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(p2)

```

The calculation of emsi still retains some of the same trends as the original NDVI, but it is now sensitive to changes in vegetation condition during the winter, spring, and fall seasons.

```{r echo=FALSE, message=FALSE, warning=FALSE}
range_dates = aoi4[1:406,]

p4 <- ggplot(range_dates[!is.na(range_dates$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-5.0,5.0), breaks = seq(-5,5,0.5)) + 
  ggtitle("Landsat 7 EMSI") + 
  xlab("Date") + 
  ylab("EMSI") + 
  geom_line(size = 0.1, color="white")

p4 <- p4 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p4)
ggsave('~/Documents/landsat7_mgio_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(p4)
```

I also show the 16-day average of NDVI over the 18 year time series (early 1984 until 2017) for all three satellites.

```{r echo=FALSE, message=FALSE, warning=FALSE}

p2 <- ggplot(aoi2[!is.na(aoi2$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-4,1.5), breaks = seq(-4,1.5,0.5)) + 
  ggtitle("Landsat 7 High Peak EMSI") + 
  xlab("Date") + 
  ylab("EMSI") + 
  geom_line(size = 0.1, color="blue")

p2 <- p2 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    legend.box.background = element_blank(),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggsave('~/Documents/landsat7_highpeak_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

plot(p2)

p3 <- ggplot(aoi1[!is.na(aoi1$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-4,1.5), breaks = seq(-4,1.5,0.5)) + 
  ggtitle("Landsat 5 High Peak EMSI") + 
  xlab("Date") + 
  ylab("EMSI") + 
  geom_line(size = 0.1)

p3 <- p3 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    legend.box.background = element_blank(),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggsave('~/Documents/landsat5_highpeak_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

plot(p3)
```

Alternately, I can plot the three time series together.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot time series
l7 = aoi2[1:406,]

p2 <- ggplot(l7, aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", high = "green", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-4,1.5), breaks = seq(-4,1.5,0.5)) +
  #geom_line(data = l5, aes(x=asdate,y=emsi), size = 0.1, color = "yellow") + 
  geom_line(data = l7, aes(x=asdate,y=emsi), size = 0.1, color = "white") +
  ggtitle("High Peak EMSI") + 
  xlab("Calendar Date") + ylab("EMSI")


p2 <- p2 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggsave('~/Documents/landsat7_highpeak_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

plot(p2)
ggplotly(p2)
```

