---
title: "EMSI LANDSAT 5,7 Pinalenos"
author: "Tyson Lee Swetnam"
date: "May 15, 2018"
output:
  pdf_document: default
  html_document: default
always_allow_html: yes
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import packages

```{r message=FALSE, warning=FALSE}
# install.packages("PerformanceAnalytics")
# install.packages("lubridate")
# install.packages("dlm")
# library(devtools)
# devtools::install_github('hadley/ggplot2')
# install_github('sinhrks/ggfortify')
# install.packages("foreign")
# install.packages("truncreg")
# install.packages("boot")
# install.packages("plotly")
```

```{r message=FALSE, warning=FALSE}
library(PerformanceAnalytics)
library(ggplot2)
library(lubridate)
library(plotly)
```

## Data

I extracted the entire time series of 16-day Landsat  NDVI data from Google Earth Engine (GEE) for an AOI on the Cienega Ranch, near the Mt Graham International Observatory (MGIO) and nearby High Peak (Mt. Graham).

In GEE I filtered the time series using the Landsat Surface Reflectance Tier 1 data.

```{r message=FALSE, warning=FALSE}
## Import the Landsat 16-day NDVI from Google Earth Engine extract

aoi4 <- read.csv("/Users/tyson_swetnam/emsi/data/landsat7_mgio_ndvi.csv")
aoi3 <- read.csv("/Users/tyson_swetnam/emsi/data/landsat5_mgio_ndvi.csv")
aoi2 <- read.csv("/Users/tyson_swetnam/emsi/data/landsat7_highpeak_ndvi.csv")
aoi1 <- read.csv("/Users/tyson_swetnam/emsi/data/landsat5_highpeak_ndvi.csv")

## Remove days with low NDVI values - likely due to cluds and snow.

aoi4$ndvi_nans <- as.numeric(as.character(aoi4$ndvi))
aoi4$ndvi_range <- ifelse(aoi4$ndvi_nans>0,aoi4$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi4$asdate <- as.Date(aoi4$date, format = "%b - %d - %Y")
aoi4$julian <- yday(aoi4$asdate)
aoi4$julian_rounded <- round((aoi4$julian/365)*52)*7

aoi3$ndvi_nans <- as.numeric(as.character(aoi3$ndvi))
aoi3$ndvi_range <- ifelse(aoi3$ndvi_nans>0.3,aoi3$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi3$asdate <- as.Date(aoi3$date, format = "%b - %d - %Y")
aoi3$julian <- yday(aoi3$asdate)
aoi3$julian_rounded <- round((aoi3$julian/365)*52)*7

aoi2$ndvi_nans <- as.numeric(as.character(aoi2$ndvi))
aoi2$ndvi_range <- ifelse(aoi2$ndvi_nans>0,aoi2$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi2$asdate <- as.Date(aoi2$date, format = "%b - %d - %Y")
aoi2$julian <- yday(aoi2$asdate)
aoi2$julian_rounded <- round((aoi2$julian/365)*52)*7

aoi1$ndvi_nans <- as.numeric(as.character(aoi1$ndvi))
aoi1$ndvi_range <- ifelse(aoi1$ndvi_nans>0,aoi1$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi1$asdate <- as.Date(aoi1$date, format = "%b - %d - %Y")
aoi1$julian <- yday(aoi1$asdate)
aoi1$julian_rounded <- round((aoi1$julian/365)*52)*7

```

Plot NDVI:
```{r message=FALSE, warning=FALSE}
# Plot examples for a range of dates
range_dates = aoi3[1:478,]
p1 <- ggplot(range_dates[!is.na(range_dates$ndvi_range),], aes(x=asdate, y=ndvi_range))+ geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 5 1984-2011 16-Day NDVI MGIO") + xlab("Date") + ylab("NDVI") + geom_line(color='white', size = 0.1) 

p1 <- p1 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggplotly(p1)
plot(p1)

ggsave('~/Documents/landsat5_mgio_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")
```

In order to make EMSI sensitive to inter-annual variation, e.g. vegetation can be more stressed than average during fall or spring dry seasons, I calculate a moving average over the Julian calendar year (1 - 365(6) days) for the time series I had available from Landsat 7.

Here I calculate EMSI during a reference period between 1984 and 1996

```{r message=FALSE, warning=FALSE}

# Predict the Julian date average of Landsat 7 NDVI using a Loess function
mean_loess7 <- predict(loess(ndvi_range ~ julian_rounded, aoi3[1:196,], span=0.1), aoi3[1:196,]$julian_rounded)
# Calculate mean NDVI for each 16-day Julian period
ndvi_mean7 <- setNames(aggregate(aoi3[1:196,]$ndvi_range, list(aoi3[1:196,]$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))

# Calculate emsi using Performance Analytics sd.multiperiod
aoi4$emsi = (aoi4$ndvi_range - mean_loess7) /  sd.multiperiod(aoi4$ndvi_range,scale=1)
aoi3$emsi = (aoi3$ndvi_range - mean_loess7) /  sd.multiperiod(aoi3$ndvi_range,scale=1)
aoi2$emsi = (aoi2$ndvi_range - mean_loess7) /  sd.multiperiod(aoi2$ndvi_range,scale=1)
aoi1$emsi = (aoi1$ndvi_range - mean_loess7) /  sd.multiperiod(aoi1$ndvi_range,scale=1)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Histograms of the time series
hist(aoi3[1:196,]$julian_rounded,breaks=52,xlab="Julian day of year", main = "Landsat 5 Histogram of Dates in Reference Dataset")
```

I used Loess function (curved blue line) to find the moving average of each period. The black line links the means of each 16-day period. 

```{r message=FALSE, warning=FALSE}
# Landsat 5 MGIO

p2 <- ggplot(aoi3[1:196,], aes(x=julian_rounded,y=ndvi_range)) + 
  geom_point(aes(color=ndvi_range)) + 
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + 
  ggtitle("Landsat 5 1984-1996 Reference 16-Day Moving Average NDVI over MGIO") + 
  xlab("Julian Date") + 
  ylab("NDVI") + 
  geom_smooth(method="loess", se=TRUE, level=0.95) 

p2 <- p2 + theme(
    line = element_line(colour = "blue"),
    title = element_text(colour = "blue"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "blue"),
    legend.text = element_text(colour = "blue"),
    axis.text = element_text(colour = "blue"),
    axis.ticks = element_line(colour = "blue"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "blue"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p2)
ggsave('~/Documents/landsat5_mgio_ndvi_mean.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(p2)

```

The calculation of emsi still retains some of the same trends as the original NDVI, but it is now sensitive to changes in vegetation condition during the winter, spring, and fall seasons.

```{r echo=FALSE, message=FALSE, warning=FALSE}
range_dates = aoi4[1:406,]

p4 <- ggplot(range_dates[!is.na(range_dates$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-5.0,5.0), breaks = seq(-5,5,0.5)) + 
  ggtitle("Landsat 7 EMSI") + 
  xlab("Date") + 
  ylab("EMSI") + 
  geom_line(size = 0.1, color="white")

p4 <- p4 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p4)
ggsave('~/Documents/landsat7_mgio_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(p4)
```

I also show the 16-day average of NDVI over the 18 year time series (early 1984 until 2017) for all three satellites.

```{r echo=FALSE, message=FALSE, warning=FALSE}

p2 <- ggplot(aoi2[!is.na(aoi2$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-4,1.5), breaks = seq(-4,1.5,0.5)) + 
  ggtitle("Landsat 7 High Peak EMSI") + 
  xlab("Date") + 
  ylab("EMSI") + 
  geom_line(size = 0.1, color="blue")

p2 <- p2 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    legend.box.background = element_blank(),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggsave('~/Documents/landsat7_highpeak_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

plot(p2)

p3 <- ggplot(aoi1[!is.na(aoi1$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-4,1.5), breaks = seq(-4,1.5,0.5)) + 
  ggtitle("Landsat 5 High Peak EMSI") + 
  xlab("Date") + 
  ylab("EMSI") + 
  geom_line(size = 0.1)

p3 <- p3 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    legend.box.background = element_blank(),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggsave('~/Documents/landsat5_highpeak_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

plot(p3)
```

Alternately, I can plot the three time series together.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot time series
l7 = aoi2[1:406,]

p2 <- ggplot(l7, aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi)) + 
  scale_colour_gradient2(low = "red", high = "green", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-4,1.5), breaks = seq(-4,1.5,0.5)) +
  #geom_line(data = l5, aes(x=asdate,y=emsi), size = 0.1, color = "yellow") + 
  geom_line(data = l7, aes(x=asdate,y=emsi), size = 0.1, color = "white") +
  ggtitle("High Peak EMSI") + 
  xlab("Calendar Date") + ylab("EMSI")


p2 <- p2 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ggsave('~/Documents/landsat7_highpeak_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

plot(p2)
ggplotly(p2)
```

