---
output: html_document
editor_options: 
  chunk_output_type: console
---
 ---
title: "Figure 4 NDVI & EMSI time series of LANDSAT 5,7,8 & Sentinel-2"
author: "Tyson Lee Swetnam"
date: "September 13, 2019"
output:
  pdf_document: default
  html_document: default
always_allow_html: yes
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import packages and data

If you havent run the script before, you may need to install some dependencies and packages. The following code block is commented out, remove leading `#` and run the code block to install packages.

```{r message=FALSE, warning=FALSE}
# Uncomment the following lines to update and install missing packages

# update.packages(checkBuilt = TRUE, ask=FALSE)
# install.packages("PerformanceAnalytics", dependencies=TRUE)
# install.packages("lubridate", dependencies=TRUE)
# install.packages("dlm", dependencies=TRUE)
# library(devtools)
# install.packages("scales", dependencies=TRUE)
# devtools::install_github('hadley/ggplot2', dependencies=TRUE)
# install_github('sinhrks/ggfortify', dependencies=TRUE)
# install.packages("foreign", dependencies=TRUE)
# install.packages("truncreg", dependencies=TRUE)
# install.packages("boot", dependencies=TRUE)
# install.packages("plotly", dependencies=TRUE)
```

## Load Packages

This code block loads the libraries needed for the graphics. 
```{r message=FALSE, warning=FALSE}
library(PerformanceAnalytics)
library(ggplot2)
library(lubridate)
library(plotly)
library(gridExtra)
```

## Create Multi-Plot ability

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Multiple plot function
#
# ggplot objects can be passed in ..., or to plotlist (as a list of ggplot objects)
# - cols:   Number of columns in layout
# - layout: A matrix specifying the layout. If present, 'cols' is ignored.
#
# If the layout is something like matrix(c(1,2,3,3), nrow=2, byrow=TRUE),
# then plot 1 will go in the upper left, 2 will go in the upper right, and
# 3 will go all the way across the bottom.
#
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  require(grid)

  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }

 if (numPlots==1) {
    print(plots[[1]])

  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

## Import Example Data

In the Git repository `/data/examples` folder I've saved data exported from the Google Earth Engine (GEE) NDVI Explorer (.js files provided in the `/javascripts` folder).

I have extracted the time series of 16-day Landsat 5,7,8 and 5-day Sentinel-2 NDVI data using their Surface Reflectance model with bit-mask flags for cloud, cloud shadow, water, or snow (Landsat), scenes with cloud <20% (Sentinel-2). Even with these filters turned on there are some pixels affected by snow or cloud contaminated pixels.

```{r message=FALSE, warning=FALSE}
## Import Landsat 5,7,8 16-day and Sentinel-2 5 day NDVI from Google Earth Engine extracted CSVs

## Okavango Thornscrub
aoi1 <- read.csv("~/emsi/data/examples/l5-okavango-shrubland-ndvi.csv")
names(aoi1) <- c("date", "ndvi")
aoi2 <- read.csv("~/emsi/data/examples/l7-okavango-shrubland-ndvi.csv")
names(aoi2) <- c("date", "ndvi")
aoi3 <- read.csv("~/emsi/data/examples/l8-okavango-shrubland-ndvi.csv")
names(aoi3) <- c("date", "ndvi")
aoi4 <- read.csv("~/emsi/data/examples/s2-okavango-shrubland-ndvi.csv")
names(aoi4) <- c("date", "ndvi")

## Okavango Swamp
aoi5 <- read.csv("~/emsi/data/examples/l5-okavango-swamp-ndvi.csv")
names(aoi5) <- c("date", "ndvi")
aoi6 <- read.csv("~/emsi/data/examples/l7-okavango-swamp-ndvi.csv")
names(aoi6) <- c("date", "ndvi")
aoi7 <- read.csv("~/emsi/data/examples/l8-okavanog-swamp-ndvi.csv")
names(aoi7) <- c("date", "ndvi")
aoi8 <- read.csv("~/emsi/data/examples/s2-okavango-swamp-ndvi.csv")
names(aoi8) <- c("date", "ndvi")

## US Mexico Grassland Mexico
aoi9 <- read.csv("~/emsi/data/examples/l5-usmex-grassmx-ndvi.csv")
names(aoi9) <- c("date", "ndvi")
aoi10 <- read.csv("~/emsi/data/examples/l7-usmex-grassmx-ndvi.csv")
names(aoi10) <- c("date", "ndvi")
aoi11 <- read.csv("~/emsi/data/examples/l8-usmex-grassmx-ndvi.csv")
names(aoi11) <- c("date", "ndvi")
aoi12 <- read.csv("~/emsi/data/examples/s2-usmex-grassmx-ndvi.csv")
names(aoi12) <- c("date", "ndvi")

## US Mexico Grassland USA
aoi13 <- read.csv("~/emsi/data/examples/l5-usmex-grassus-ndvi.csv")
names(aoi13) <- c("date", "ndvi")
aoi14 <- read.csv("~/emsi/data/examples/l7-usmex-grassus-ndvi.csv")
names(aoi14) <- c("date", "ndvi")
aoi15 <- read.csv("~/emsi/data/examples/l8-usmex-grassus-ndvi.csv")
names(aoi15) <- c("date", "ndvi")
aoi16 <- read.csv("~/emsi/data/examples/s2-usmex-grassus-ndvi.csv")
names(aoi16) <- c("date", "ndvi")

## Acre Brazil Agricultural Areas
aoi17 <- read.csv("~/emsi/data/examples/l5-acre-ag-ndvi.csv")
names(aoi17) <- c("date", "ndvi")
aoi18 <- read.csv("~/emsi/data/examples/l7-acre-ag-ndvi.csv")
names(aoi18) <- c("date", "ndvi")
aoi19 <- read.csv("~/emsi/data/examples/l8-acre-ag-ndvi.csv")
names(aoi19) <- c("date", "ndvi")
aoi20 <- read.csv("~/emsi/data/examples/s2-acre-ag-ndvi.csv")
names(aoi20) <- c("date", "ndvi")

## Acre Brazil Rainforest
aoi21 <- read.csv("~/emsi/data/examples/l5-acre-rainforest-ndvi.csv")
names(aoi21) <- c("date", "ndvi")
aoi22 <- read.csv("~/emsi/data/examples/l7-acre-rainforest-ndvi.csv")
names(aoi22) <- c("date", "ndvi")
aoi23 <- read.csv("~/emsi/data/examples/l8-acre-rainforest-ndvi.csv")
names(aoi23) <- c("date", "ndvi")
aoi24 <- read.csv("~/emsi/data/examples/s2-acre-rainforest-ndvi.csv")
names(aoi24) <- c("date", "ndvi")

## Yakutia Taiga
aoi25 <- read.csv("~/emsi/data/examples/l5-yakutia-taiga-ndvi.csv")
names(aoi25) <- c("date", "ndvi")
aoi26 <- read.csv("~/emsi/data/examples/l7-yakutia-taiga-ndvi.csv")
names(aoi26) <- c("date", "ndvi")
aoi27 <- read.csv("~/emsi/data/examples/l8-yakutia-taiga-ndvi.csv")
names(aoi27) <- c("date", "ndvi")
aoi28 <- read.csv("~/emsi/data/examples/s2-yakutia-taiga-ndvi.csv")
names(aoi28) <- c("date", "ndvi")

## Yakutia Alaas
aoi29 <- read.csv("~/emsi/data/examples/l5-yakutia-alaas-ndvi.csv")
names(aoi29) <- c("date", "ndvi")
aoi30 <- read.csv("~/emsi/data/examples/l7-yakutia-alaas-ndvi.csv")
names(aoi30) <- c("date", "ndvi")
aoi31 <- read.csv("~/emsi/data/examples/l8-yakutia-alaas-ndvi.csv")
names(aoi31) <- c("date", "ndvi")
aoi32 <- read.csv("~/emsi/data/examples/s2-yakutia-alaas-ndvi.csv")
names(aoi32) <- c("date", "ndvi")
```

## Calculate Julian dates from calendar

Create a function to set the Julian (1-366) dates from the data, also flags any dates with no data.
```{r message=FALSE, warning=FALSE}
# Create Functions for calculating Julian Dates 
aoi.date_julian <-  function(df){
	df$ndvi_nans <- as.numeric(as.character(df$ndvi))
	df$ndvi_range <- ifelse(df$ndvi_nans>0.05,df$ndvi_nans,NA)
	df$ndvi_range <- ifelse(df$ndvi_nans<0.95,df$ndvi_nans,NA)
	df$asdate <- as.Date(df$date, format = "%b %d, %Y")
	df$julian <- yday(df$asdate)
	df$julian_rounded <- round((df$julian/365)*52)*7
	df
}
```

## Execute Functions recursively through all of the `aoi`

Runs the Julian function on the 32 example data sets.
```{r message=FALSE, warning=FALSE}
aoi1 <- aoi.date_julian(aoi1)
aoi2 <- aoi.date_julian(aoi2)
aoi3 <- aoi.date_julian(aoi3)
aoi4 <- aoi.date_julian(aoi4)
aoi5 <- aoi.date_julian(aoi5)
aoi6 <- aoi.date_julian(aoi6)
aoi7 <- aoi.date_julian(aoi7)
aoi8 <- aoi.date_julian(aoi8)
aoi9 <- aoi.date_julian(aoi9)
aoi10 <- aoi.date_julian(aoi10)
aoi11 <- aoi.date_julian(aoi11)
aoi12 <- aoi.date_julian(aoi12)
aoi13 <- aoi.date_julian(aoi13)
aoi14 <- aoi.date_julian(aoi14)
aoi15 <- aoi.date_julian(aoi15)
aoi16 <- aoi.date_julian(aoi16)
aoi17 <- aoi.date_julian(aoi17)
aoi18 <- aoi.date_julian(aoi18)
aoi19 <- aoi.date_julian(aoi19)
aoi20 <- aoi.date_julian(aoi20)
aoi21 <- aoi.date_julian(aoi21)
aoi22 <- aoi.date_julian(aoi22)
aoi23 <- aoi.date_julian(aoi23)
aoi24 <- aoi.date_julian(aoi24)
aoi25 <- aoi.date_julian(aoi25)
aoi26 <- aoi.date_julian(aoi26)
aoi27 <- aoi.date_julian(aoi27)
aoi28 <- aoi.date_julian(aoi28)
aoi29 <- aoi.date_julian(aoi29)
aoi30 <- aoi.date_julian(aoi30)
aoi31 <- aoi.date_julian(aoi31)
aoi32 <- aoi.date_julian(aoi32)
```

## Calculate the loess smoothed mean NDVI for each Julian date and EMSI

Create a function that calculates the EMSI from the NDVI Julian date time series. 
```{r message=FALSE, warning=FALSE}
# Calculate EMSI for each area of interest
aoi.emsi <-  function(df){
	mean_loess <- predict(loess(ndvi_range ~ julian_rounded, df), df$julian_rounded)
	df$emsi = (df$ndvi_range - mean_loess) / sd.multiperiod(df$ndvi_range,scale=1)
	df
}
```

Runs the EMSI function on the 32 example data sets.
```{r message=FALSE, warning=FALSE}
aoi1 <- aoi.emsi(aoi1)
aoi2 <- aoi.emsi(aoi2)
aoi3 <- aoi.emsi(aoi3)
aoi4 <- aoi.emsi(aoi4)
aoi5 <- aoi.emsi(aoi5)
aoi6 <- aoi.emsi(aoi6)
aoi7 <- aoi.emsi(aoi7)
aoi8 <- aoi.emsi(aoi8)
aoi9 <- aoi.emsi(aoi9)
aoi10 <- aoi.emsi(aoi10)
aoi11 <- aoi.emsi(aoi11)
aoi12 <- aoi.emsi(aoi12)
aoi13 <- aoi.emsi(aoi13)
aoi14 <- aoi.emsi(aoi14)
aoi15 <- aoi.emsi(aoi15)
aoi16 <- aoi.emsi(aoi16)
aoi17 <- aoi.emsi(aoi17)
aoi18 <- aoi.emsi(aoi18)
aoi19 <- aoi.emsi(aoi19)
aoi20 <- aoi.emsi(aoi20)
aoi21 <- aoi.emsi(aoi21)
aoi22 <- aoi.emsi(aoi22)
aoi23 <- aoi.emsi(aoi23)
aoi24 <- aoi.emsi(aoi24)
aoi25 <- aoi.emsi(aoi25)
aoi26 <- aoi.emsi(aoi26)
aoi27 <- aoi.emsi(aoi27)
aoi28 <- aoi.emsi(aoi28)
aoi29 <- aoi.emsi(aoi29)
aoi30 <- aoi.emsi(aoi30)
aoi31 <- aoi.emsi(aoi31)
aoi32 <- aoi.emsi(aoi32)
```

### Merge Landsat 5, 7, 8, and Sentinel-2 time series

This code block creates a single data.frame with all of the NDVI time series. Currently, it is commented out to limit the effect of Sentinel-2 cloud artifacts in the mean NDVI calculation.
```{r message=FALSE, warning=FALSE}
#all1 <- rbind(aoi1,aoi2,aoi3,aoi4)
#all2 <- rbind(aoi5,aoi6,aoi7,aoi8)
#all3 <- rbind(aoi9,aoi10,aoi11,aoi12)
#all4 <- rbind(aoi13,aoi14,aoi15,aoi16)
#all5 <- rbind(aoi17,aoi18,aoi19,aoi20)
#all6 <- rbind(aoi21,aoi22,aoi23,aoi24)
#all7 <- rbind(aoi25,aoi26,aoi27,aoi28)
#all8 <- rbind(aoi29,aoi30,aoi31,aoi32)
```

### Merge Landsat 5, 7, 8 time series
```{r message=FALSE, warning=FALSE}
all1 <- rbind(aoi1,aoi2,aoi3)
all2 <- rbind(aoi5,aoi6,aoi7)
all3 <- rbind(aoi9,aoi10,aoi11)
all4 <- rbind(aoi13,aoi14,aoi15)
all5 <- rbind(aoi17,aoi18,aoi19)
all6 <- rbind(aoi21,aoi22,aoi23)
all7 <- rbind(aoi25,aoi26,aoi27)
all8 <- rbind(aoi29,aoi30,aoi31)
```

## Calculate the Loess regression spine for the mean NDVI and 

We use a `loess` regression to show the running mean for the NDVI and EMSI time series.
### Okavango 
```{r message=FALSE, warning=FALSE}
# Calculate mean NDVI for Julian period in Shrubland
ndvi_mean_shrubland <- c(NA)
ndvi_mean_shrubland$l5 <- setNames(aggregate(aoi1$ndvi_range, list(aoi1$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_shrubland$l7 <- setNames(aggregate(aoi2$ndvi_range, list(aoi2$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_shrubland$l8 <- setNames(aggregate(aoi3$ndvi_range, list(aoi3$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_shrubland$s2 <- setNames(aggregate(aoi4$ndvi_range, list(aoi4$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_shrubland$all <- setNames(aggregate(all1$ndvi_range, list(all1$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
```

```{r message=FALSE, warning=FALSE}
# Calculate mean NDVI for Julian period in Swamp
ndvi_mean_swamp <- c(NA)
ndvi_mean_swamp$l5 <- setNames(aggregate(aoi5$ndvi_range, list(aoi5$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_swamp$l7 <- setNames(aggregate(aoi6$ndvi_range, list(aoi6$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_swamp$l8 <- setNames(aggregate(aoi7$ndvi_range, list(aoi7$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_swamp$s2 <- setNames(aggregate(aoi8$ndvi_range, list(aoi8$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_swamp$all <- setNames(aggregate(all2$ndvi_range, list(all2$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
```

### US Mexico
```{r message=FALSE, warning=FALSE}
# Calculate mean NDVI for Julian period in Mexico
ndvi_mean_grassmx <- c(NA)
ndvi_mean_grassmx$l5 <- setNames(aggregate(aoi9$ndvi_range, list(aoi9$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_grassmx$l7 <- setNames(aggregate(aoi10$ndvi_range, list(aoi10$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_grassmx$l8 <- setNames(aggregate(aoi11$ndvi_range, list(aoi11$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_grassmx$s2 <- setNames(aggregate(aoi12$ndvi_range, list(aoi12$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_grassmx$all <- setNames(aggregate(all3$ndvi_range, list(all3$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
```

```{r message=FALSE, warning=FALSE}
# Calculate mean NDVI for Julian period in USA
ndvi_mean_grassus <- c(NA)
ndvi_mean_grassus$l5 <- setNames(aggregate(aoi13$ndvi_range, list(aoi13$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_grassus$l7 <- setNames(aggregate(aoi14$ndvi_range, list(aoi14$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_grassus$l8 <- setNames(aggregate(aoi15$ndvi_range, list(aoi15$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_grassus$s2 <- setNames(aggregate(aoi16$ndvi_range, list(aoi16$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_grassus$all <- setNames(aggregate(all4$ndvi_range, list(all4$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
```

### Acre
```{r message=FALSE, warning=FALSE}
# Calculate mean NDVI for Julian period in Agricultural
ndvi_mean_ag <- c(NA)
ndvi_mean_ag$l5 <- setNames(aggregate(aoi17$ndvi_range, list(aoi17$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_ag$l7 <- setNames(aggregate(aoi18$ndvi_range, list(aoi18$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_ag$l8 <- setNames(aggregate(aoi19$ndvi_range, list(aoi19$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_ag$s2 <- setNames(aggregate(aoi20$ndvi_range, list(aoi20$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_ag$all <- setNames(aggregate(all5$ndvi_range, list(all5$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
```

```{r message=FALSE, warning=FALSE}
# Calculate mean NDVI for Julian period in Rainforest
ndvi_mean_rainforest <- c(NA)
ndvi_mean_rainforest$l5 <- setNames(aggregate(aoi21$ndvi_range, list(aoi21$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_rainforest$l7 <- setNames(aggregate(aoi22$ndvi_range, list(aoi22$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_rainforest$l8 <- setNames(aggregate(aoi23$ndvi_range, list(aoi23$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_rainforest$s2 <- setNames(aggregate(aoi24$ndvi_range, list(aoi24$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_rainforest$all <- setNames(aggregate(all6$ndvi_range, list(all6$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
```

### Yakutia
```{r message=FALSE, warning=FALSE}
# Calculate mean NDVI for Julian period in Taiga
ndvi_mean_taiga <- c(NA)
ndvi_mean_taiga$l5 <- setNames(aggregate(aoi25$ndvi_range, list(aoi25$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_taiga$l7 <- setNames(aggregate(aoi26$ndvi_range, list(aoi26$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_taiga$l8 <- setNames(aggregate(aoi27$ndvi_range, list(aoi27$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_taiga$s2 <- setNames(aggregate(aoi28$ndvi_range, list(aoi28$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_taiga$all <- setNames(aggregate(all7$ndvi_range, list(all7$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
```

```{r message=FALSE, warning=FALSE}
# Calculate mean NDVI for Julian period in Alaas
ndvi_mean_alaas <- c(NA)
ndvi_mean_alaas$l5 <- setNames(aggregate(aoi29$ndvi_range, list(aoi29$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_alaas$l7 <- setNames(aggregate(aoi30$ndvi_range, list(aoi30$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_alaas$l8 <- setNames(aggregate(aoi31$ndvi_range, list(aoi31$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_alaas$s2 <- setNames(aggregate(aoi32$ndvi_range, list(aoi32$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
ndvi_mean_alaas$all <- setNames(aggregate(all8$ndvi_range, list(all8$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
```

## Plot Scene Frequency Histograms

Histograms show the frequency of scences that are available to derive mean values.

To plot additional histograms, change the `aoi` to another example area.
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Histograms of the four time series and all together
hist(aoi1$julian_rounded,breaks=52,xlab="Julian day of year", main = "Landsat 5 Histogram of Dates in Shrubland")
hist(aoi2$julian_rounded,breaks=52,xlab="Julian day of year", main = "Landsat 7 Histogram of Dates in Shrubland")
hist(aoi3$julian_rounded,breaks=52,xlab="Julian day of year", main = "Landsat 8 Histogram of Dates in Shrubland")
hist(aoi4$julian_rounded,breaks=52,xlab="Julian day of year", main = "Sentinel-2 Histogram of Dates in Shrubland")
hist(all1$julian_rounded,breaks=52,xlab="Julian day of year", main = "All Histogram of Dates in Shrubland")
```

# Plot Time Series 
Create sets of time series plots for the Julian dates

### Okavango Thornscrub with Legend
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Okavango Thornscrub
plot_ndvi_shrubland <- ggplot() +
  xlim(1,365) + ylim(0.0,0.75) +
  geom_point(data = aoi1[!is.na(aoi1$ndvi_range),], aes(x=julian,y=ndvi_range, color = "Landsat 5"), size = 2.0) + 
  geom_smooth(data = aoi1[!is.na(aoi1$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.5, size=0.5, color = "red") +
  geom_point(data = aoi2[!is.na(aoi2$ndvi_range),], aes(x=julian,y=ndvi_range, color = "Landsat 7"), size = 2.0) +
  geom_smooth(data = aoi2[!is.na(aoi2$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.5, size=0.5, color = "green") +
  geom_point(data = aoi3[!is.na(aoi3$ndvi_range),], aes(x=julian,y=ndvi_range, color = "Landsat 8"), size = 2.0) +
  geom_smooth(data = aoi3[!is.na(aoi3$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.5, size=0.5, color = "blue") +
  geom_point(data = aoi4[!is.na(aoi4$ndvi_range),], aes(x=julian,y=ndvi_range, color = "Sentinel-2"), size = 2.0) +
  geom_smooth(data = aoi4[!is.na(aoi4$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.5, size=0.5, color = "orange") +
  geom_smooth(data = all1[!is.na(aoi1$ndvi_range),], aes(x=julian,y=ndvi_range, color = "Landsat-Series"), method="loess", se = FALSE, na.rm = FALSE, span = 0.5, size=1) + 
  scale_color_manual(name = "EOS Platform", values = c("red","green","blue","black","orange")) + # Colors are placed in order from above series
  ggtitle("Okavango Thornscrub") + 
  xlab("Julian Day") + ylab("NDVI") + theme(legend.key=element_blank()) 

plot_ndvi_shrubland <- plot_ndvi_shrubland + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "black"),
    legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_ndvi_shrubland)

ggsave('~/emsi/rmd/examples/fig_4_ndvi_okavango_shrubland.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_ndvi_shrubland)

```
### Okavango Thornscrub
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Okavango Thornscrub
plot_ndvi_shrubland <- ggplot() +
   xlim(1,365) + ylim(0.0,0.75) +
  geom_point(data = aoi1[!is.na(aoi1$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "red") + 
  geom_smooth(data = aoi1[!is.na(aoi1$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.5, color = "red", size=0.5) +
  geom_point(data = aoi2[!is.na(aoi2$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "green") +
  geom_smooth(data = aoi2[!is.na(aoi2$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.5, color = "green", size=0.5) +
  geom_point(data = aoi3[!is.na(aoi3$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "blue") +
  geom_smooth(data = aoi3[!is.na(aoi3$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.5, color = "blue", size=0.5) +
  geom_point(data = aoi4[!is.na(aoi4$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "orange") +
  geom_smooth(data = aoi4[!is.na(aoi4$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.5, color = "orange", size=0.5) +
  geom_smooth(data = all1[!is.na(aoi1$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.5, color = "black", size=1) +
  ggtitle("Okavango Thornscrub") + 
  xlab("Julian Day") + ylab("NDVI") + theme(legend.key=element_blank()) 

plot_ndvi_shrubland <- plot_ndvi_shrubland + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "back"),
    legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_ndvi_shrubland)
ggsave('~/emsi/rmd/examples/ndvi_okavango_shrubland.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_ndvi_shrubland)

```

### Okavango Swamp
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Okavango Swamp
plot_ndvi_swamp <- ggplot() +
   xlim(1,365) + ylim(0.0,0.75) +
  geom_point(data = aoi5[!is.na(aoi5$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "red") + 
  geom_smooth(data = aoi5[!is.na(aoi5$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi6[!is.na(aoi6$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "green") +
  geom_smooth(data = aoi6[!is.na(aoi6$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi7[!is.na(aoi7$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "blue") +
  geom_smooth(data = aoi7[!is.na(aoi7$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi8[!is.na(aoi8$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "orange") +
  geom_smooth(data = aoi8[!is.na(aoi8$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all2[!is.na(aoi2$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "black", size=1) +
  ggtitle("Okavango Swamp") + 
  xlab("Julian Day") + ylab("NDVI")

plot_ndvi_swamp <- plot_ndvi_swamp + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "back"),
    legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_ndvi_swamp)
ggsave('~/emsi/rmd/examples/ndvi_okavango_swamp.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_ndvi_swamp)

```

### US-Mex Border Mexico
```{r echo=FALSE, message=FALSE, warning=FALSE}
# US Mexico Grassland Mexico
plot_ndvi_grassmx <- ggplot() +
  xlim(1,365) + ylim(0.0,1.0) +
  geom_point(data = aoi9[!is.na(aoi9$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "red") + 
  geom_smooth(data = aoi9[!is.na(aoi9$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi10[!is.na(aoi10$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "green") +
  geom_smooth(data = aoi10[!is.na(aoi10$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi11[!is.na(aoi11$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "blue") +
  geom_smooth(data = aoi11[!is.na(aoi11$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi12[!is.na(aoi12$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "orange") +
  geom_smooth(data = aoi12[!is.na(aoi12$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all3[!is.na(aoi3$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "black", size=1) +
  ggtitle("Mexico Grassland") + 
  xlab("Julian Day") + ylab("NDVI")

plot_ndvi_grassmx <- plot_ndvi_grassmx + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "back"),
    legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_ndvi_grassmx)
ggsave('~/emsi/rmd/examples/ndvi_usmex_grassmx.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_ndvi_grassmx)

```
### US-Mex Border USA
```{r echo=FALSE, message=FALSE, warning=FALSE}
# US Mexico Grassland USA
plot_ndvi_grassus <- ggplot() +
  xlim(1,365) + ylim(0.0,1.0) +
  geom_point(data = aoi13[!is.na(aoi13$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "red") + 
  geom_smooth(data = aoi13[!is.na(aoi13$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi14[!is.na(aoi14$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "green") +
  geom_smooth(data = aoi14[!is.na(aoi14$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi15[!is.na(aoi15$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "blue") +
  geom_smooth(data = aoi15[!is.na(aoi15$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi16[!is.na(aoi16$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "orange") +
  geom_smooth(data = aoi16[!is.na(aoi16$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all4[!is.na(aoi4$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "black", size=1) +
  ggtitle("USA Grassland") + 
  xlab("Julian Day") + ylab("NDVI")

plot_ndvi_grassus <- plot_ndvi_grassus + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "back"),
    legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_ndvi_grassus)
ggsave('~/emsi/rmd/examples/ndvi_usmex_grassus.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_ndvi_grassus)
```
### Acre Agriculture
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Acre Agriculture
plot_ndvi_acreag <- ggplot() +
  xlim(1,365) + ylim(0.0,1.0) +
  geom_point(data = aoi17[!is.na(aoi17$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "red") + 
  geom_smooth(data = aoi17[!is.na(aoi17$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi18[!is.na(aoi18$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "green") +
  geom_smooth(data = aoi18[!is.na(aoi18$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi19[!is.na(aoi19$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "blue") +
  geom_smooth(data = aoi19[!is.na(aoi19$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi20[!is.na(aoi20$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "orange") +
  geom_smooth(data = aoi20[!is.na(aoi20$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all5[!is.na(aoi5$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "black", size=1) +
  ggtitle("Acre Brazil Agriculture") + 
  xlab("Julian Day") + ylab("NDVI")

plot_ndvi_acreag <- plot_ndvi_acreag + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "back"),
    legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_ndvi_acreag)
ggsave('~/emsi/rmd/examples/ndvi_acreag.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_ndvi_acreag)
```
### Acre Rainforest
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Acre Rainforest
plot_ndvi_acrerainforest <- ggplot() +
  xlim(1,365) + ylim(0.0,1.0) +
  geom_point(data = aoi21[!is.na(aoi21$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "red") + 
  geom_smooth(data = aoi21[!is.na(aoi21$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi22[!is.na(aoi22$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "green") +
  geom_smooth(data = aoi22[!is.na(aoi22$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi23[!is.na(aoi23$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "blue") +
  geom_smooth(data = aoi23[!is.na(aoi23$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi24[!is.na(aoi24$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "orange") +
  geom_smooth(data = aoi24[!is.na(aoi24$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all6[!is.na(aoi6$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "black", size=1) +
  ggtitle("Acre Brazil Rainforest") + 
  xlab("Julian Day") + ylab("NDVI")

plot_ndvi_acrerainforest <- plot_ndvi_acrerainforest + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "back"),
    legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_ndvi_acrerainforest)
ggsave('~/emsi/rmd/examples/ndvi_acrerainforest.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_ndvi_acrerainforest)
```
### Yakutia Taiga
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Yakutia Taiga
plot_ndvi_yakutiataiga <- ggplot() +
  xlim(1,365) + ylim(0.0,1.0) +
  geom_point(data = aoi25[!is.na(aoi25$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "red") + 
  geom_smooth(data = aoi25[!is.na(aoi25$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi26[!is.na(aoi26$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "green") +
  geom_smooth(data = aoi26[!is.na(aoi26$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi27[!is.na(aoi27$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "blue") +
  geom_smooth(data = aoi27[!is.na(aoi27$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi28[!is.na(aoi28$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "orange") +
  geom_smooth(data = aoi28[!is.na(aoi28$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all7[!is.na(aoi7$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "black", size=1) +
  ggtitle("Yakutia Taiga") + 
  xlab("Julian Day") + ylab("NDVI")

plot_ndvi_yakutiataiga <- plot_ndvi_yakutiataiga + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "back"),
    legend.text = element_text(colour = "black"),
    legend.position = "bottom",
    legend.box = "horizontal",
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
) 

plot(plot_ndvi_yakutiataiga)

ggsave('~/emsi/rmd/examples/ndvi_yakutiataiga.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_ndvi_yakutiataiga)
```
### Yakutia Alaas
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Yakutia Taiga
plot_ndvi_yakutiaalaas <- ggplot() +
  xlim(1,365) + ylim(0.0,1.0) +
  geom_point(data = aoi29[!is.na(aoi29$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "red") + 
  geom_smooth(data = aoi29[!is.na(aoi29$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi30[!is.na(aoi30$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "green") +
  geom_smooth(data = aoi30[!is.na(aoi30$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi31[!is.na(aoi31$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "blue") +
  geom_smooth(data = aoi31[!is.na(aoi31$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi32[!is.na(aoi32$ndvi_range),], aes(x=julian,y=ndvi_range), size = 2.0, color = "orange") +
  geom_smooth(data = aoi32[!is.na(aoi32$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all8[!is.na(aoi8$ndvi_range),], aes(x=julian,y=ndvi_range), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "black", size=1) +
  ggtitle("Yakutia Alaas") + 
  xlab("Julian Day") + ylab("NDVI")

plot_ndvi_yakutiaalaas <- plot_ndvi_yakutiaalaas + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "back"),
    legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_ndvi_yakutiaalaas)
ggsave('~/emsi/rmd/examples/ndvi_yakutiaalaas.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_ndvi_yakutiaalaas)
```
### Okavango Thornscrub

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Okavango Thornscrub
plot_emsi_shrubland <- ggplot() +
  xlim(1,365) + ylim(-2,2) +
  geom_point(data = aoi1[!is.na(aoi1$emsi),], aes(x=julian,y=emsi, color = "Landsat 5"), size = 2.0) + 
  geom_smooth(data = aoi1[!is.na(aoi1$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 1, size=0.5, color = "red") +
  geom_point(data = aoi2[!is.na(aoi2$emsi),], aes(x=julian,y=emsi, color = "Landsat 7"), size = 2.0) +
  geom_smooth(data = aoi2[!is.na(aoi2$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 1, size=0.5, color = "green") +
  geom_point(data = aoi3[!is.na(aoi3$emsi),], aes(x=julian,y=emsi, color = "Landsat 8"), size = 2.0) +
  geom_smooth(data = aoi3[!is.na(aoi3$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 1, size=0.5, color = "blue") +
  geom_point(data = aoi4[!is.na(aoi4$emsi),], aes(x=julian,y=emsi, color = "Sentinel-2"), size = 2.0) +
  geom_smooth(data = aoi4[!is.na(aoi4$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 1, size=0.5, color = "orange") +
  geom_smooth(data = all1[!is.na(aoi1$emsi),], aes(x=julian,y=emsi, color = "Landsat-Series"), method="loess", se = FALSE, na.rm = FALSE, span = 1, size=1) + 
  scale_color_manual(name = "EOS Platform", values = c("red","green","blue","black","orange")) + # Colors are placed in order from above series
   ggtitle("Okavango Thornscrub") + 
  xlab("Julian Day") + ylab("EMSI") + theme(legend.key=element_blank()) 

plot_emsi_shrubland <- plot_emsi_shrubland + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "black"),
    legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_emsi_shrubland)

ggsave('~/emsi/rmd/examples/fig_5_emsi_okavango_shrubland.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_emsi_shrubland)

```

### Okavango Swamp
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Okavango Swamp
plot_emsi_swamp <- ggplot() +
  xlim(1,365) + ylim(-3,3) +
  geom_point(data = aoi5[!is.na(aoi5$emsi),], aes(x=julian,y=emsi), size = 2.0, color = "red") + 
  geom_smooth(data = aoi5[!is.na(aoi5$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi6[!is.na(aoi6$emsi),], aes(x=julian,y=emsi), size = 2.0, color="green") +
  geom_smooth(data = aoi6[!is.na(aoi6$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi7[!is.na(aoi7$emsi),], aes(x=julian,y=emsi), size = 2.0, color="blue") +
  geom_smooth(data = aoi7[!is.na(aoi7$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi8[!is.na(aoi8$emsi),], aes(x=julian,y=emsi), size = 2.0, color="orange") +
  geom_smooth(data = aoi8[!is.na(aoi8$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all2[!is.na(aoi2$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, size=1, color="black") + 
  scale_color_manual(values = c("red","green","blue","black","orange")) + # Colors are given an alphabetical order from above
  ggtitle("Okavango Swamp") + 
  xlab("Julian Day") + ylab("EMSI")

plot_emsi_swamp <- plot_emsi_swamp + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_emsi_swamp)

ggsave('~/emsi/rmd/examples/emsi_okavango_swamp.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_emsi_swamp)

```

### Mexico Grassland
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Mexico Grassland
plot_emsi_grassmx <- ggplot() +
  xlim(1,365) + ylim(-3,3) +
  geom_point(data = aoi9[!is.na(aoi9$emsi),], aes(x=julian,y=emsi), size = 2.0, color = "red") + 
  geom_smooth(data = aoi9[!is.na(aoi9$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi10[!is.na(aoi10$emsi),], aes(x=julian,y=emsi), size = 2.0, color="green") +
  geom_smooth(data = aoi10[!is.na(aoi10$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi11[!is.na(aoi11$emsi),], aes(x=julian,y=emsi), size = 2.0, color="blue") +
  geom_smooth(data = aoi11[!is.na(aoi11$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi12[!is.na(aoi12$emsi),], aes(x=julian,y=emsi), size = 2.0, color="orange") +
  geom_smooth(data = aoi12[!is.na(aoi12$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all3[!is.na(aoi3$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, size=1, color="black") + 
  scale_color_manual(values = c("red","green","blue","black","orange")) + # Colors are given an alphabetical order from above
  ggtitle("Mexico Grassland") + 
  xlab("Julian Day") + ylab("EMSI")

plot_emsi_grassmx <- plot_emsi_grassmx + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_emsi_grassmx)

ggsave('~/emsi/rmd/examples/emsi_mx_grass.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_emsi_grassmx)

```

### USA Grassland
```{r echo=FALSE, message=FALSE, warning=FALSE}
# USA Grassland
plot_emsi_grassus <- ggplot() +
  xlim(1,365) + ylim(-3,3) +
  geom_point(data = aoi13[!is.na(aoi13$emsi),], aes(x=julian,y=emsi), size = 2.0, color = "red") + 
  geom_smooth(data = aoi13[!is.na(aoi13$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi14[!is.na(aoi14$emsi),], aes(x=julian,y=emsi), size = 2.0, color="green") +
  geom_smooth(data = aoi14[!is.na(aoi14$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi15[!is.na(aoi15$emsi),], aes(x=julian,y=emsi), size = 2.0, color="blue") +
  geom_smooth(data = aoi15[!is.na(aoi15$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi16[!is.na(aoi16$emsi),], aes(x=julian,y=emsi), size = 2.0, color="orange") +
  geom_smooth(data = aoi16[!is.na(aoi16$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all4[!is.na(aoi4$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, size=1, color="black") + 
  scale_color_manual(values = c("red","green","blue","black","orange")) + # Colors are given an alphabetical order from above
  ggtitle("USA Grassland") + 
  xlab("Julian Day") + ylab("EMSI")

plot_emsi_grassus <- plot_emsi_grassus + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_emsi_grassus)

ggsave('~/emsi/rmd/examples/emsi_us_grass.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_emsi_grassus)

```

### Acre Agriculture
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Acre Agriculture
plot_emsi_acreag <- ggplot() +
  xlim(1,365) + ylim(-3,3) +
  geom_point(data = aoi17[!is.na(aoi17$emsi),], aes(x=julian,y=emsi), size = 2.0, color = "red") + 
  geom_smooth(data = aoi17[!is.na(aoi17$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi18[!is.na(aoi18$emsi),], aes(x=julian,y=emsi), size = 2.0, color="green") +
  geom_smooth(data = aoi18[!is.na(aoi18$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi19[!is.na(aoi19$emsi),], aes(x=julian,y=emsi), size = 2.0, color="blue") +
  geom_smooth(data = aoi19[!is.na(aoi19$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi20[!is.na(aoi20$emsi),], aes(x=julian,y=emsi), size = 2.0, color="orange") +
  geom_smooth(data = aoi20[!is.na(aoi20$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all5[!is.na(aoi5$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, size=1, color="black") + 
  scale_color_manual(values = c("red","green","blue","black","orange")) + # Colors are given an alphabetical order from above
  ggtitle("Acre Agriculture") + 
  xlab("Julian Day") + ylab("EMSI")

plot_emsi_acreag <- plot_emsi_acreag + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_emsi_acreag)

ggsave('~/emsi/rmd/examples/emsi_acre_ag.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_emsi_acreag)

```

### Acre Rainforest
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Acre Rainforest
plot_emsi_acrerainforest <- ggplot() +
  xlim(1,365) + ylim(-3,3) +
  geom_point(data = aoi21[!is.na(aoi21$emsi),], aes(x=julian,y=emsi), size = 2.0, color = "red") + 
  geom_smooth(data = aoi21[!is.na(aoi21$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi22[!is.na(aoi22$emsi),], aes(x=julian,y=emsi), size = 2.0, color="green") +
  geom_smooth(data = aoi22[!is.na(aoi22$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi19[!is.na(aoi23$emsi),], aes(x=julian,y=emsi), size = 2.0, color="blue") +
  geom_smooth(data = aoi23[!is.na(aoi23$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi24[!is.na(aoi24$emsi),], aes(x=julian,y=emsi), size = 2.0, color="orange") +
  geom_smooth(data = aoi24[!is.na(aoi24$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all6[!is.na(aoi6$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, size=1, color="black") + 
  scale_color_manual(values = c("red","green","blue","black","orange")) + # Colors are given an alphabetical order from above
  ggtitle("Acre Rainforest") + 
  xlab("Julian Day") + ylab("EMSI")

plot_emsi_acrerainforest <- plot_emsi_acrerainforest + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_emsi_acrerainforest)

ggsave('~/emsi/rmd/examples/emsi_acre_rain.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_emsi_acrerainforest)

```


### Yakutia Taiga
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Yakutia Taiga
plot_emsi_taiga <- ggplot() +
  xlim(1,365) + ylim(-3,3) +
  geom_point(data = aoi25[!is.na(aoi25$emsi),], aes(x=julian,y=emsi), size = 2.0, color = "red") + 
  geom_smooth(data = aoi25[!is.na(aoi25$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi26[!is.na(aoi26$emsi),], aes(x=julian,y=emsi), size = 2.0, color="green") +
  geom_smooth(data = aoi26[!is.na(aoi26$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi27[!is.na(aoi27$emsi),], aes(x=julian,y=emsi), size = 2.0, color="blue") +
  geom_smooth(data = aoi27[!is.na(aoi27$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi28[!is.na(aoi28$emsi),], aes(x=julian,y=emsi), size = 2.0, color="orange") +
  geom_smooth(data = aoi28[!is.na(aoi28$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all7[!is.na(aoi7$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, size=1, color="black") + 
  scale_color_manual(values = c("red","green","blue","black","orange")) + # Colors are given an alphabetical order from above
  ggtitle("Yakutia Taiga") + 
  xlab("Julian Day") + ylab("EMSI")

plot_emsi_taiga <- plot_emsi_taiga + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_emsi_taiga)

ggsave('~/emsi/rmd/examples/emsi_taiga.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_emsi_taiga)

```

### Yakutia Alaas
```{r echo=FALSE, message=FALSE, warning=FALSE}
# Yakutia Taiga
plot_emsi_alaas <- ggplot() +
  xlim(1,365) + ylim(-3,3) +
  geom_point(data = aoi29[!is.na(aoi29$emsi),], aes(x=julian,y=emsi), size = 2.0, color = "red") + 
  geom_smooth(data = aoi29[!is.na(aoi29$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "red", size=0.5) +
  geom_point(data = aoi30[!is.na(aoi30$emsi),], aes(x=julian,y=emsi), size = 2.0, color="green") +
  geom_smooth(data = aoi30[!is.na(aoi30$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "green", size=0.5) +
  geom_point(data = aoi31[!is.na(aoi31$emsi),], aes(x=julian,y=emsi), size = 2.0, color="blue") +
  geom_smooth(data = aoi31[!is.na(aoi31$emsi),], aes(x=julian,y=emsi),method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "blue", size=0.5) +
  geom_point(data = aoi32[!is.na(aoi32$emsi),], aes(x=julian,y=emsi), size = 2.0, color="orange") +
  geom_smooth(data = aoi32[!is.na(aoi32$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, color = "orange", size=0.5) +
  geom_smooth(data = all8[!is.na(aoi8$emsi),], aes(x=julian,y=emsi), method="loess", se = FALSE, na.rm = FALSE, span = 0.2, size=1, color="black") + 
  scale_color_manual(values = c("red","green","blue","black","orange")) + # Colors are given an alphabetical order from above
  ggtitle("Yakutia Alaas") + 
  xlab("Julian Day") + ylab("EMSI")

plot_emsi_alaas <- plot_emsi_alaas + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_line(colour = "gray"), 
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(plot_emsi_alaas)

ggsave('~/emsi/rmd/examples/emsi_alaas.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(plot_emsi_alaas)

```

## Plot all Julian NDVI and EMSI
```{r echo=FALSE, message=FALSE, warning=FALSE}
## EMSI and NDVI for Julian Dates
m1 <- multiplot(plot_emsi_shrubland, plot_ndvi_shrubland, plot_emsi_swamp, plot_ndvi_swamp, plot_emsi_grassmx, plot_ndvi_grassmx, plot_emsi_grassus, plot_ndvi_grassus, cols=2)

m2 <- multiplot(plot_emsi_acreag, plot_ndvi_acreag, plot_emsi_acrerainforest, plot_ndvi_acrerainforest, plot_emsi_taiga, plot_ndvi_yakutiataiga, plot_emsi_alaas, plot_ndvi_yakutiaalaas, cols=2)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot just NDVI for the Julian dates
m1 <- multiplot(plot_ndvi_yakutiataiga, plot_ndvi_yakutiaalaas, plot_ndvi_acreag, plot_ndvi_acrerainforest, cols=2)
m2 <- multiplot(plot_ndvi_grassmx, plot_ndvi_grassus, plot_ndvi_shrubland, plot_ndvi_swamp, cols=2)
```
# NDVI Time Series

These plot use the calendar dates to create NDVI time series
### Okavango Thornscrub
```{r echo=FALSE, message=FALSE, warning=FALSE}
ndvi_shrubland <- ggplot(all1[!is.na(all1$ndvi),], aes(x=asdate,y=ndvi)) + 
  geom_point(aes(color=ndvi), size=2.0) + 
  scale_colour_gradient2(low = "cyan", mid = "green", high = "darkgreen", midpoint = 0.5, limits=c(0,0.95)) + 
  scale_y_continuous(limits = c(0,0.8), breaks = seq(0,1,0.1)) + geom_smooth(data = all1[!is.na(all1$ndvi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +
  ggtitle("Okavango Thornscrub") + 
  xlab("Date") + 
  ylab("NDVI") 

ndvi_shrubland <- ndvi_shrubland + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

ndvi_shrubland + scale_fill_discrete(guide = guide_legend(reverse=TRUE))

ggsave('~/emsi/rmd/examples/okavango_shrubland_landsat578s2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(ndvi_shrubland)
```
### Okavango Swamp
```{r echo=FALSE, message=FALSE, warning=FALSE}
ndvi_swamp <- ggplot(all2[!is.na(all2$ndvi),], aes(x=asdate,y=ndvi)) + 
  geom_point(aes(color=ndvi), size=2.0) + 
   scale_colour_gradient2(low = "cyan", mid = "green", high = "darkgreen", midpoint = 0.5, limits=c(0,0.95)) + 
  scale_y_continuous(limits = c(0,0.8), breaks = seq(0,1,0.1)) + geom_smooth(data = all2[!is.na(all2$ndvi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +  ggtitle("Okavango Swamp") + 
  xlab("Date") + 
  ylab("NDVI") 

ndvi_swamp <- ndvi_swamp + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(ndvi_swamp)
ggsave('~/emsi/rmd/examples/okavango_swamp_landsat578s2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(ndvi_swamp)
```
### Mexico Grassland
```{r echo=FALSE, message=FALSE, warning=FALSE}
ndvi_grassmx <- ggplot(all3[!is.na(all3$ndvi),], aes(x=asdate,y=ndvi)) + 
  geom_point(aes(color=ndvi), size=2.0) + 
  scale_colour_gradient2(low = "cyan", mid = "green", high = "darkgreen", midpoint = 0.5, limits=c(0,0.95)) + 
  scale_y_continuous(limits = c(0,0.7), breaks = seq(0,1,0.1)) + geom_smooth(data = all3[!is.na(all3$ndvi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +  ggtitle("Mexico Grassland") + 
  xlab("Date") + 
  ylab("NDVI") 

ndvi_grassmx <- ndvi_grassmx + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(ndvi_grassmx)
ggsave('~/emsi/rmd/examples/usmex_grassmx_landsat578s2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(ndvi_grassmx)
```
### USA Grassland
```{r echo=FALSE, message=FALSE, warning=FALSE}
ndvi_grassus <- ggplot(all4[!is.na(all4$ndvi),], aes(x=asdate,y=ndvi)) + 
  geom_point(aes(color=ndvi), size=2.0) + 
  scale_colour_gradient2(low = "cyan", mid = "green", high = "darkgreen", midpoint = 0.5, limits=c(0,0.95)) + 
  scale_y_continuous(limits = c(0,0.7), breaks = seq(0,1,0.1)) + geom_smooth(data = all4[!is.na(all4$ndvi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +  ggtitle("USA Grassland") + 
  xlab("Date") + 
  ylab("NDVI") 

ndvi_grassus <- ndvi_grassus + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(ndvi_grassus)
ggsave('~/emsi/rmd/examples/usmex_grassus_landsat578s2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(ndvi_grassus)
```
### Acre Agriculture
```{r echo=FALSE, message=FALSE, warning=FALSE}
ndvi_acreag <- ggplot(all5[!is.na(all5$ndvi),], aes(x=asdate,y=ndvi)) + 
  geom_point(aes(color=ndvi), size=2.0) + 
  scale_colour_gradient2(low = "cyan", mid = "green", high = "darkgreen", midpoint = 0.5, limits=c(0,0.95)) + 
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.1)) + geom_smooth(data = all5[!is.na(all5$ndvi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +  ggtitle("Acre Agriculture") + 
  xlab("Date") + 
  ylab("NDVI") 

ndvi_acreag <- ndvi_acreag + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(ndvi_acreag)
ggsave('~/emsi/rmd/examples/usmex_acreag_landsat578s2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(ndvi_acreag)
```
### Acre Rainforest
```{r echo=FALSE, message=FALSE, warning=FALSE}
ndvi_acrerainforest <- ggplot(all6[!is.na(all6$ndvi),], aes(x=asdate,y=ndvi)) + 
  geom_point(aes(color=ndvi), size=2.0) + 
  scale_colour_gradient2(low = "cyan", mid = "green", high = "darkgreen", midpoint = 0.5, limits=c(0,0.95)) + 
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.1)) + geom_smooth(data = all6[!is.na(all6$ndvi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +  ggtitle("Acre Rainforest") + 
  xlab("Date") + 
  ylab("NDVI") 

ndvi_acrerainforest <- ndvi_acrerainforest + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(ndvi_acrerainforest)
ggsave('~/emsi/rmd/examples/usmex_acrerainforest_landsat578s2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(ndvi_acrerainforest)
```
### Yakutia Taiga
```{r echo=FALSE, message=FALSE, warning=FALSE}
ndvi_taiga <- ggplot(all7[!is.na(all7$ndvi),], aes(x=asdate,y=ndvi)) + 
  geom_point(aes(color=ndvi), size=2.0) + 
  scale_colour_gradient2(low = "cyan", mid = "green", high = "darkgreen", midpoint = 0.5, limits=c(0,0.95)) + 
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.1)) + geom_smooth(data = all7[!is.na(all7$ndvi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +  ggtitle("Yakutia Taiga") + 
  xlab("Date") + 
  ylab("NDVI") 

ndvi_taiga <- ndvi_taiga + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
   #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(ndvi_taiga)
ggsave('~/emsi/rmd/examples/taiga_landsat578s2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(ndvi_taiga)
```
### Yakutia Alaas
```{r echo=FALSE, message=FALSE, warning=FALSE}
ndvi_alaas <- ggplot(all8[!is.na(all8$ndvi),], aes(x=asdate,y=ndvi)) + 
  geom_point(aes(color=ndvi), size=2.0) + 
  scale_colour_gradient2(low = "cyan", mid = "green", high = "darkgreen", midpoint = 0.5, limits=c(0,0.95)) + 
  scale_y_continuous(limits = c(0,1), breaks = seq(0,1,0.1)) + geom_smooth(data = all8[!is.na(all8$ndvi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) + 
  ggtitle("Yakutia Alaas") + 
  xlab("Date") + 
  ylab("NDVI") 

ndvi_alaas <- ndvi_alaas + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(ndvi_alaas)
ggsave('~/emsi/rmd/examples/alaas_landsat578s2_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(ndvi_alaas)
```

# EMSI Time Series
These plot use the calendar dates to create EVI time series
### Okavango Thornscrub
```{r echo=FALSE, message=FALSE, warning=FALSE}
emsi_shrubland <- ggplot(all1[!is.na(all1$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi), size=2.0) + geom_line(aes(x=asdate,y=emsi), size= 0.1) +
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.0, limits=c(-2.3,2.3)) + 
  scale_y_continuous(limits = c(-2.3,2.3), breaks = seq(-2.5,2.5,0.5)) + 
  #geom_smooth(data = all1[!is.na(all2$emsi),], method="auto", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=0.5, alpha=0.2) +
  ggtitle("Okavango Thornscrub") + 
  xlab("Date") + 
  ylab("EMSI") 

emsi_shrubland <- emsi_shrubland + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

emsi_shrubland + scale_fill_discrete(guide = guide_legend(reverse=TRUE))

ggsave('~/emsi/rmd/examples/okavango_shrubland_landsat578s2_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(emsi_shrubland)
```
### Okavango Swamp
```{r echo=FALSE, message=FALSE, warning=FALSE}
emsi_swamp <- ggplot(all2[!is.na(all2$emsi),], aes(x=asdate,y=emsi)) + 
   geom_point(aes(color=emsi), size=2.0) + geom_line(aes(x=asdate,y=emsi), size= 0.1) +
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.0, limits=c(-2.3,2.3)) + 
  scale_y_continuous(limits = c(-2.3,2.3), breaks = seq(-2.5,2.5,0.5)) + 
  #geom_smooth(data = all2[!is.na(all2$emsi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +
  ggtitle("Okavango Swamp") + 
  xlab("Date") + 
  ylab("EMSI") 

emsi_swamp <- emsi_swamp + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(emsi_swamp)
ggsave('~/emsi/rmd/examples/okavango_swamp_landsat578s2_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(emsi_swamp)
```
### Mexico Grassland
```{r echo=FALSE, message=FALSE, warning=FALSE}
emsi_grassmx <- ggplot(all3[!is.na(all3$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi), size=2.0) + 
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.0, limits=c(-2.3,2.3)) + 
  scale_y_continuous(limits = c(-2.5,2.5), breaks = seq(-2.5,2.5,0.5)) + geom_smooth(data = all3[!is.na(all3$emsi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +
  ggtitle("Mexico Grassland") + 
  xlab("Date") + 
  ylab("EMSI") 

emsi_grassmx <- emsi_grassmx + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(emsi_grassmx)
ggsave('~/emsi/rmd/examples/usmex_grassmx_landsat578s2_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(emsi_grassmx)
```
### USA Grassland
```{r echo=FALSE, message=FALSE, warning=FALSE}
emsi_grassus <- ggplot(all4[!is.na(all4$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi), size=2.0) + 
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.0, limits=c(-2.3,2.3)) + 
  scale_y_continuous(limits = c(-2.5,2.5), breaks = seq(-2.5,2.5,0.5)) + geom_smooth(data = all4[!is.na(all4$emsi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +
  ggtitle("USA Grassland") + 
  xlab("Date") + 
  ylab("EMSI") 

emsi_grassus <- emsi_grassus + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(emsi_grassus)
ggsave('~/emsi/rmd/examples/usmex_grassus_landsat578s2_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(emsi_grassus)
```
### Acre Agriculture
```{r echo=FALSE, message=FALSE, warning=FALSE}
emsi_acreag <- ggplot(all5[!is.na(all5$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi), size=2.0) + 
  scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.0, limits=c(-2.3,2.3)) + 
  scale_y_continuous(limits = c(-2.5,2.5), breaks = seq(-2.5,2.5,0.5)) + geom_smooth(data = all5[!is.na(all5$emsi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +
  ggtitle("Acre Agriculture") + 
  xlab("Date") + 
  ylab("EMSI") 

emsi_acreag <- emsi_acreag + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(emsi_acreag)
ggsave('~/emsi/rmd/examples/usmex_acreag_landsat578s2_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(emsi_acreag)
```
### Acre Rainforest
```{r echo=FALSE, message=FALSE, warning=FALSE}
emsi_acrerainforest <- ggplot(all6[!is.na(all6$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi), size=2.0) + 
    scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.0, limits=c(-2.3,2.3)) + 
  scale_y_continuous(limits = c(-2.5,2.5), breaks = seq(-2.5,2.5,0.5)) + geom_smooth(data = all6[!is.na(all6$emsi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +
  ggtitle("Acre Rainforest") + 
  xlab("Date") + 
  ylab("EMSI") 

emsi_acrerainforest <- emsi_acrerainforest + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(emsi_acrerainforest)
ggsave('~/emsi/rmd/examples/usmex_acrerainforest_landsat578s2_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(emsi_acrerainforest)
```
### Yakutia Taiga
```{r echo=FALSE, message=FALSE, warning=FALSE}
emsi_taiga <- ggplot(all7[!is.na(all7$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi), size=2.0) + 
    scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.0, limits=c(-2.3,2.3)) + 
  scale_y_continuous(limits = c(-2.5,2.5), breaks = seq(-2.5,2.5,0.5)) + geom_smooth(data = all7[!is.na(all7$emsi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +
  ggtitle("Yakutia Taiga") + 
  xlab("Date") + 
  ylab("EMSI") 

emsi_taiga <- emsi_taiga + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
   #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(emsi_taiga)
ggsave('~/emsi/rmd/examples/taiga_landsat578s2_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(emsi_taiga)
```
### Yakutia Alaas
```{r echo=FALSE, message=FALSE, warning=FALSE}
emsi_alaas <- ggplot(all8[!is.na(all8$emsi),], aes(x=asdate,y=emsi)) + 
  geom_point(aes(color=emsi), size=2.0) + 
    scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.0, limits=c(-2.3,2.3)) + 
  scale_y_continuous(limits = c(-2.5,2.5), breaks = seq(-2.5,2.5,0.5)) + geom_smooth(data = all8[!is.na(all8$emsi),], method="loess", se = FALSE, na.rm = TRUE, span = 0.025, color = "black", size=1.0) +
  ggtitle("Yakutia Alaas") + 
  xlab("Date") + 
  ylab("EMSI") 

emsi_alaas <- emsi_alaas + theme(
    line = element_line(colour = "black"),
    title = element_text(colour = "black"),
    #legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    #legend.box.background = element_rect(fill = "transparent"),
    #legend.title = element_text(colour = "black"),
    #legend.text = element_text(colour = "black"),
    axis.text = element_text(colour = "black"),
    axis.ticks = element_line(colour = "black"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "black"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(emsi_alaas)
ggsave('~/emsi/rmd/examples/alaas_landsat578s2_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(emsi_alaas)
```
### Plot all Calendar NDVI and EMSI 
```{r}
m1 <- multiplot(emsi_shrubland, ndvi_shrubland, emsi_swamp, ndvi_swamp, emsi_grassmx, ndvi_grassmx, emsi_grassus, ndvi_grassus, cols=2)

m2 <- multiplot(emsi_acreag, ndvi_acreag, emsi_acrerainforest, ndvi_acrerainforest, emsi_taiga, ndvi_taiga, emsi_alaas, ndvi_alaas, cols=2)
```


```{r}
m1 <- multiplot(emsi_shrubland, ndvi_shrubland,cols=1)

```
