---
title: "emsi LANDSAT 5,7,8"
author: "Tyson Lee Swetnam"
date: "May 10, 2018"
output:
  pdf_document: default
  html_document: default
always_allow_html: yes
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import packages

```{r message=FALSE, warning=FALSE}
# install.packages("PerformanceAnalytics")
# install.packages("lubridate")
# install.packages("dlm")
# library(devtools)
# devtools::install_github('hadley/ggplot2')
# install_github('sinhrks/ggfortify')
# install.packages("foreign")
# install.packages("truncreg")
# install.packages("boot")
# install.packages("plotly")
```

```{r message=FALSE, warning=FALSE}
library(PerformanceAnalytics)
library(ggplot2)
library(lubridate)
library(plotly)
```

## Import Data

I extracted the entire time series of 16-day Landsat  NDVI data from Google Earth Engine (GEE) for RAWS weather sites using a ~500 meter boxed area of interest. I chose the larger sample area over the point observation of an indvidual pixel to gain more confidence in trends of the 16-day variation over time. 

In GEE I filtered the data using the Landsat [LEDAPS product](https://landsat.usgs.gov/sites/default/files/documents/ledaps_product_guide.pdf) for clouds, cloud shadows, adjacent clouds, and snow. 

Even with these filters turned on there are some scenes with snow or cloud contaminated pixels, these generally drop the NDVI value below a minimum of ~0.4 for mature pine forest areas.

Here I am using data from around the Rincon RAWS site east of Tucson, Arizona. 

```{r message=FALSE, warning=FALSE}
## Import the Landsat 16-day NDVI from Google Earth Engine extract
aoi5 <- read.csv("/Users/tyson_swetnam/Downloads/landsat5_rincon.csv")
aoi7 <- read.csv("/Users/tyson_swetnam/Downloads/landsat7_rincon.csv")
aoi8 <- read.csv("/Users/tyson_swetnam/Downloads/landsat8_rincon.csv")

## Remove days with low NDVI values - likely due to cluds and snow.

# Landsat 5
aoi5$ndvi_nans <- as.numeric(as.character(aoi5$ndvi))
aoi5$ndvi_range <- ifelse(aoi5$ndvi_nans>0.4,aoi5$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi5$asdate <- as.Date(aoi5$date, format = "%d - %b - %y")
aoi5$julian <- yday(aoi5$asdate)
aoi5$julian_rounded <- round((aoi5$julian/365)*52)*7

# Landsat 7
aoi7$ndvi_nans <- as.numeric(as.character(aoi7$ndvi))
aoi7$ndvi_range <- ifelse(aoi7$ndvi_nans>0.4,aoi7$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi7$asdate <- as.Date(aoi7$date, format = "%d - %b - %y")
aoi7$julian <- yday(aoi7$asdate)
aoi7$julian_rounded <- round((aoi7$julian/365)*52)*7

# Landsat 8
aoi8$ndvi_nans <- as.numeric(as.character(aoi8$ndvi))
aoi8$ndvi_range <- ifelse(aoi8$ndvi_nans>0.4,aoi8$ndvi_nans,NA)
## Convert the dates to an R readable format
aoi8$asdate <- as.Date(aoi8$date, format = "%d - %b - %y")
aoi8$julian <- yday(aoi8$asdate)
aoi8$julian_rounded <- round((aoi8$julian/365)*52)*7
```

The problem I found with using a global z-score: `emsi = obs_ndvi - mean_ndvi / stdDev_ndvi` was that a single mean and standard deviation resulted in an emsi closely resembles the NDVI time series from which it is generated. 

```{r message=FALSE, warning=FALSE}
## Calculate NDVI average (old emsi calculation)
# Landsat 5
ndvi_mean_original5 <- mean(aoi5$ndvi_range, na.rm=TRUE)
ndvi_stdDev_original5 <- sd(aoi5$ndvi_range,  na.rm=TRUE)
aoi5$emsi_original <- (aoi5$ndvi_range - ndvi_mean_original5) / ndvi_stdDev_original5
# Landsat 7
ndvi_mean_original7 <- mean(aoi7$ndvi_range, na.rm=TRUE)
ndvi_stdDev_original7 <- sd(aoi7$ndvi_range,  na.rm=TRUE)
aoi7$emsi_original <- (aoi7$ndvi_range - ndvi_mean_original7) / ndvi_stdDev_original7
# Landsat 8
ndvi_mean_original8 <- mean(aoi8$ndvi_range, na.rm=TRUE)
ndvi_stdDev_original8 <- sd(aoi8$ndvi_range,  na.rm=TRUE)
aoi8$emsi_original <- (aoi8$ndvi_range - ndvi_mean_original8) / ndvi_stdDev_original8
```

Plot NDVI:
```{r message=FALSE, warning=FALSE}
# Plot examples for a range of dates
range_dates = aoi5[1:50,]
p1 <- ggplot(range_dates[!is.na(range_dates$ndvi_range),], aes(x=asdate, y=ndvi_range))+ geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.45) + ggtitle("Landsat 5 1984-1986 16-Day NDVI") + xlab("Date") + ylab("NDVI") + geom_line(color='white', size = 0.1) 

p1 <- p1 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)
ggplotly(p1)
plot(p1)
ggsave('~/Documents/landsat5_example_ndvi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")
```
Plot emsi:
```{r message=FALSE, warning=FALSE}
range_dates = aoi5[1:50,]
p2 <- ggplot(range_dates[!is.na(range_dates$emsi_original),], aes(x=asdate, y=emsi_original))+ geom_point(aes(color=emsi_original)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.0) + ggtitle("Landsat 5 16-Day emsi Global Calculation") + xlab("Date") + ylab("EMSI Global") + geom_line(color='white', size = 0.1)

p2 <- p2 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)
ggplotly(p2)
plot(p2)
ggsave('~/Documents/landsat5_example_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")
```

The global calculation of FSMI does not account for the variation in vegetation phenology, i.e. vegetation is  robustly green following even weak seasonal rainfalls and the emsi is positive. 

In order to make emsi more sensitive to inter-annual variation, e.g. vegetation can be more stressed than average during fall or spring dry seasons, I calculated a moving average over the Julian calendar year (1 - 365(6) days) for the time series I had available from each satellite platform (Landsats 5, 7, 8). 

There are 981 scenes for Landsat 5, 768 scenese for Landsat 7, and 265 scenes for Landsat 8 in this dataset.

```{r message=FALSE, warning=FALSE}
# Predict the Julian date average of Landsat 5 NDVI using a Loess function
mean_loess5 <- predict(loess(ndvi_range ~ julian_rounded, aoi5), aoi5$julian_rounded)
# Calculate mean NDVI for each 16-day Julian period
ndvi_mean5 <- setNames(aggregate(aoi5$ndvi_range, list(aoi5$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
# Calculate emsi using Performance Analytics sd.multiperiod
aoi5$emsi = (aoi5$ndvi_range - mean_loess5) /  sd.multiperiod(aoi5$ndvi_range,scale=1)

# Predict the Julian date average of Landsat 7 NDVI using a Loess function
mean_loess7 <- predict(loess(ndvi_range ~ julian_rounded, aoi7), aoi7$julian_rounded)
# Calculate mean NDVI for each 16-day Julian period
ndvi_mean7 <- setNames(aggregate(aoi7$ndvi_range, list(aoi7$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
# Calculate emsi using Performance Analytics sd.multiperiod
aoi7$emsi = (aoi7$ndvi_range - mean_loess7) /  sd.multiperiod(aoi7$ndvi_range,scale=1)

# Predict the Julian date average of Landsat 8 NDVI using a Loess function
mean_loess8 <- predict(loess(ndvi_range ~ julian_rounded, aoi8), aoi8$julian_rounded)
# Calculate mean NDVI for each 16-day Julian period
ndvi_mean8 <- setNames(aggregate(aoi8$ndvi_range, list(aoi8$julian_rounded), mean, na.rm=TRUE, na.action=NULL), c("julian", "ndvi_mean"))
# Calculate emsi using Performance Analytics sd.multiperiod
aoi8$emsi = (aoi8$ndvi_range - mean_loess8) /  sd.multiperiod(aoi8$ndvi_range,scale=1)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Histograms of the three time series
hist(aoi5$julian_rounded,breaks=52,xlab="Julian day of year", main = "Landsat 5 Histogram of Dates in Dataset")
hist(aoi7$julian_rounded,breaks=52,xlab="Julian day of year", main = "Landsat 7 Histogram of Dates in Dataset")
hist(aoi8$julian_rounded,breaks=52,xlab="Julian day of year", main = "Landsat 8 Histogram of Dates in Dataset")
```

I used Loess function (curved blue line) to find the moving average of each period. The black line links the means of each 16-day period. 

```{r message=FALSE, warning=FALSE}
# Landsat 5
p6 <- ggplot(aoi5, aes(x=julian_rounded,y=ndvi_range)) + geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 5 1984-2014 16-Day Moving Average NDVI") + xlab("Julian Date") + ylab("NDVI") + geom_smooth(method="loess", se=TRUE, fullrange=TRUE, level=0.95) + geom_line(colour = 'white', data=ndvi_mean5, aes(x=julian,y=ndvi_mean)) + geom_point(data=ndvi_mean5, color = 'white', aes(x=julian,y=ndvi_mean))

p6 <- p6 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p6)
ggsave('~/Documents/landsat5_ndvi_mean.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(p6)

# Landsat 7
p7 <- ggplot(aoi7, aes(x=julian_rounded,y=ndvi_range)) + geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 7 1999-2017 16-Day Moving Average NDVI") + xlab("Julian Date") + ylab("NDVI") + geom_smooth(method="loess", se=TRUE, fullrange=TRUE, level=0.95) + geom_line(data=ndvi_mean7, aes(x=julian,y=ndvi_mean))+ geom_point(data=ndvi_mean7, aes(x=julian,y=ndvi_mean))

ggplotly(p7)

# Landsat 8
p8 <- ggplot(aoi8, aes(x=julian_rounded,y=ndvi_range)) + geom_point(aes(color=ndvi_range)) + scale_colour_gradient2(low = "red", mid = "green", high = "blue", midpoint = 0.5) + ggtitle("Landsat 8 2014-2017 16-Day Moving Average NDVI") + xlab("Julian Date") + ylab("NDVI") + geom_smooth(method="loess", se=TRUE, fullrange=TRUE, level=0.95) + geom_line(data=ndvi_mean8, aes(x=julian,y=ndvi_mean)) + geom_point(data=ndvi_mean8, aes(x=julian,y=ndvi_mean))

ggplotly(p8)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}

# Plot time series together
l5 = aoi5[!is.na(aoi5$ndvi_range),]
l7 = aoi7[!is.na(aoi7$ndvi_range),]
l8 = aoi8[!is.na(aoi8$ndvi_range),]

p3 <- ggplot() +
  geom_point(data = l5, aes(x=julian,y=ndvi_range), size = 0.1, color = "red") + 
  geom_line(data=ndvi_mean5, aes(x=julian,y=ndvi_mean), color = "red") +
  geom_point(data = l7, aes(x=julian,y=ndvi_range), size = 0.1, color = "green") +
  geom_line(data=ndvi_mean7, aes(x=julian,y=ndvi_mean), color = "green") +
  geom_point(data = l8, aes(x=julian,y=ndvi_range), size = 0.1, color = "blue") + 
  geom_line(data=ndvi_mean8, aes(x=julian,y=ndvi_mean), color = "blue") +
  ggtitle("NDVI for Landsats 5,7,8") + 
  xlab("Julian Day") + ylab("NDVI")

p3 <- p3 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p3)
ggsave('~/Documents/landsat578_ndvi_mean.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(p3)

```

The time series averages of the three Landsat satellites (red = Landsat 5, green = Landsat 7, blue = Landsat 8) shows how the 16-day NDVI average has changed over the last 33 years.

The new calculation of emsi still retains some of the same trends as the original emsi, but it is now sensitive to changes in vegetation condition during the winter, spring, and fall seasons.

```{r echo=FALSE, message=FALSE, warning=FALSE}
range_dates = aoi5[1:50,]

p4 <- ggplot(range_dates[!is.na(range_dates$emsi),], aes(x=asdate,y=emsi)) + geom_point(aes(color=emsi)) + scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + scale_y_continuous(limits = c(-1.3,1.3), breaks = seq(-1.3,1.3,0.5)) + ggtitle("Landsat 5 EMSI") + xlab("Date") + ylab("EMSI") + geom_line(size = 0.1)

p4 <- p4 + theme(
    line = element_line(colour = "white"),
    title = element_text(colour = "white"),
    legend.background = element_rect(fill = "transparent"), # get rid of legend bg
    legend.box.background = element_rect(fill = "transparent"),
    legend.title = element_text(colour = "white"),
    legend.text = element_text(colour = "white"),
    axis.text = element_text(colour = "white"),
    axis.ticks = element_line(colour = "white"),
    panel.background = element_rect(fill = "transparent",colour = NA), # or theme_blank()
    panel.grid.minor = element_line(colour = "white"), 
    panel.grid.major = element_blank(),
    plot.background = element_rect(fill = "transparent",colour = NA)
)

plot(p4)
ggsave('~/Documents/landsat5_emsi.png', width = 8, height = 4.5, dpi = 300, bg = "transparent")

ggplotly(p4)
```

I also show the 16-day average of NDVI over the 18 year time series (early 1984 until 2017) for all three satellites.

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aoi5[!is.na(aoi5$emsi),], aes(x=asdate,y=emsi)) + geom_point(aes(color=emsi)) + scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + scale_y_continuous(limits = c(-1.3,1.3), breaks = seq(-1.3,1.3,0.1)) + ggtitle("Landsat 5 EMSI") + xlab("Date") + ylab("EMSI") + geom_line(size = 0.1)

ggplot(aoi7[!is.na(aoi7$emsi),], aes(x=asdate,y=emsi)) + geom_point(aes(color=emsi)) + scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + scale_y_continuous(limits = c(-1.3,1.3), breaks = seq(-1.3,1.3,0.1)) + ggtitle("Landsat 7 EMSI") + xlab("Date") + ylab("EMSI") + geom_line(size = 0.1)

ggplot(aoi8[!is.na(aoi8$emsi),], aes(x=asdate,y=emsi)) + geom_point(aes(color=emsi)) + scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + scale_y_continuous(limits = c(-1.3,1.3), breaks = seq(-1.3,1.3,0.1)) + ggtitle("Landsat 8 EMSI") + xlab("Date") + ylab("EMSI") + geom_line(size = 0.1)


```

Alternately, I can plot the three time series together.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# Plot time series together
l5 = aoi5[!is.na(aoi5$emsi),]
l7 = aoi7[!is.na(aoi7$emsi),]
l8 = aoi8[!is.na(aoi8$emsi),]

p5 <- ggplot() +
  geom_line(data = l5, aes(x=asdate,y=emsi), size = 0.1, color = "red") + 
  geom_line(data = l7, aes(x=asdate,y=emsi), size = 0.1, color = "green") +
  geom_line(data = l8, aes(x=asdate,y=emsi), size = 0.1, color = "blue") + 
  scale_colour_gradient2(low = "red", mid = "green" , high = "blue", midpoint = 0.0) + 
  scale_y_continuous(limits = c(-1.3,1.3), breaks = seq(-1.3,1.3,0.1)) + 
  ggtitle("EMSI of Landsat 5,7,8") + 
  xlab("Calendar Date") + ylab("EMSI")

ggplotly(p5)

```
