---
title: "Okavango EMSI Raster Calculation"
author: "Tyson Lee Swetnam"
date: "2021-03-03"
output: html_document
---

# Setup

## Linux Dependencies

If you are running this for the first time, you may need to install additional geospatial dependencies

```{bash message=FALSE, warning=FALSE}
# sudo add-apt-repository ppa:ubuntugis
# sudo apt-get update
# sudo apt-get install libgdal-dev libgeos-dev libproj-dev
# sudo apt-get install libudunits2-dev
```

## Install missing R Libraries

The following code block executes when the notebook is opened. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("dplyr")) {
   install.packages("dplyr")
   library(dplyr)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}  
if (!require("grid")) {
   install.packages("grid")
   library(grid)
} 
if (!require("lattice")) {
   install.packages("lattice")
   library(lattice)
}
if (!require("leaflet")) {
   install.packages("leaflet")
   library(leaflet)
}
if (!require("lubridate")) {
   install.packages("lubridate")
   library(lubridate)
}
if (!require("magrittr")) {
   install.packages("magrittr")
   library(magrittr)
}
if (!require("raster")) {
   install.packages("raster")
   library(raster)  
}
if (!require("rasterVis")) {
   install.packages("rasterVis")
   library(rasterVis)   
}
if (!require("RColorBrewer")) {
   install.packages("RColorBrewer")
   library(RColorBrewer)
}
if (!require("reshape")) {
   install.packages("reshape")
   library(reshape)
}
if (!require("rgdal")) {
   install.packages("rgdal")
   library(rgdal)   
}  
if (!require("remotes")) {
   install.packages("remotes")
   library(remotes)   
}  
if (!require("PerformanceAnalytics")) {
   install.packages("PerformanceAnalytics")
   library(PerformanceAnalytics)
}
if (!require("scales")) {
   install.packages("scales")
   library(scales)
}
if (!require("sf")) {
   install.packages("sf")
   library(sf)   
}
if (!require("stringr")) {
   install.packages("stringr")
   library(stringr)
}
if (!require("viridis")) {
   install.packages("viridis")
   library(viridis)
}
```

# Transfer datasets from CyVerse

This command uses `wget` to download data into the container.

The CyVerse iRODS data store also supports [WebDav](https://data.cyverse.org); we are hosting the extracted raster data in a folder with anonymous public read-only settings.

All of the landsat data are ~22.8 GB in size, and will take ~18 minutes to transfer within the CyVerse workbench for all four sites.

A subset of the collection will take less time, here we're taking the `/okavango` collection. 

```{bash message=FALSE, warning=FALSE}
# uncomment to run
 cd ~/emsi/data/
time wget -r -nH --cut-dirs=7 --no-parent -l8 --reject="index.html*" https://data.cyverse.org/dav-anon/iplant/home/tswetnam/emsi/data/collections/okavango
```

# Import Raster from data directory and check the header
```{r message=FALSE, warning=FALSE}
# Test for raster metadata with GDALinfo
GDALinfo("~/emsi/data/collections/okavango/gee-collection-okavango-landsat8/LC08_175073_20171026.tif")
```

## Import Raster time series for Landsats 5,7,8
```{r message=FALSE, warning=FALSE}
# Load all rasters in okavango-landsat5
setwd("~/emsi/data/collections/okavango/gee-collection-okavango-landsat5/")
rlist5=list.files(getwd(), pattern=".tif$", full.names=FALSE)
for(i in rlist5) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) } 

# Load all rasters in okavango-landsat7
setwd("~/emsi/data/collections/okavango/gee-collection-okavango-landsat7/")
rlist7=list.files(getwd(), pattern="tif$", full.names=FALSE)
for(i in rlist7) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) }

# Load all rasters in okavango-landsat8
setwd("~/emsi/data/collections/okavango/gee-collection-okavango-landsat8/")
rlist8=list.files(getwd(), pattern="tif$", full.names=FALSE)
for(i in rlist8) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) }
```

```{r}
list_l5 <- ls(pattern="LT05", all.names = TRUE) 
dates_l5 = as.Date(str_sub(list_l5, -8 ,-1), format="%Y%m%d")

list_l7 <- ls(pattern="LE07", all.names = TRUE) 
dates_l7 = as.Date(str_sub(list_l7, -8 ,-1), format="%Y%m%d")

list_l8 <- ls(pattern="LC08", all.names = TRUE) 
dates_l8 = as.Date(str_sub(list_l8, -8 ,-1), format="%Y%m%d")

list_03 <- ls(pattern = "03", all.names = TRUE)
list_03
```

# March (peak greenness) EMSI calc prep
```{r}
# Landsat 5,7,8 March
lall_03 <- brick(LT05_175073_19870330, 
                      LT05_175073_19890303, 
                      LT05_175073_19890319, 
                      LT05_175073_19900322, 
                      LT05_175073_19910309, 
                      LT05_175073_19920327, 
                      LT05_175073_19930314, 
                      LT05_175073_19940317, 
                      LT05_175073_19950304, 
                      LT05_175073_19960306, 
                      LT05_175073_19960322, 
                      LT05_175073_19970309, 
                      LT05_175073_20040312, 
                      LT05_175073_20080307, 
                      LT05_175073_20090326,
                      LE07_175073_20010312, 
                      LE07_175073_20010328, 
                      LE07_175073_20020331, 
                      LE07_175073_20030318, 
                      LE07_175073_20050307, 
                      LE07_175073_20060326, 
                      LE07_175073_20070313, 
                      LE07_175073_20080331, 
                      LE07_175073_20090318, 
                      LE07_175073_20100321, 
                      LE07_175073_20130313,
                      LC08_175073_20140308, 
                      LC08_175073_20140324, 
                      LC08_175073_20150311,
                      LE07_175073_20150319,
                      LC08_175073_20150327,
                      LC08_175073_20160329, 
                      LC08_175073_20170316, 
                      LE07_175073_20170324, 
                      LE07_175073_20180327,
                      LC08_175073_20180319)

#       LC08_175074_20140308, 
#                      LC08_175074_20140324, 
#                      LC08_175074_20150311, 
#                      LC08_175074_20150327, 
#                      LC08_175074_20160329, 
#                      LC08_175074_20170316,
#                      LC08_175074_20180319, 

#       LC08_127015_20140308, 
#                      LC08_127015_20140324, 
#                      LC08_127015_20150311, 
#                     LC08_127015_20160313, 
#                      LC08_127015_20170316, 
#                      LC08_127015_20180303, 
#                      LC08_127015_20180319, 
#                      LC08_127016_20140308, 
#                      LC08_127016_20140324, 
#                      LC08_127016_20150311, 
#                      LC08_127016_20160313, 
#                      LC08_127016_20160329, 
#                      LC08_127016_20180303, 
#                      LC08_127016_20180319, 
#                      LC08_128015_20140331, 
#                      LC08_128015_20160304, 
#                      LC08_128015_20170307, 
#                      LC08_128015_20180310, 
#                      LC08_128015_20180326,
#                      LC08_129015_20140306, 
#                      LC08_129015_20140322, 
#                      LC08_129015_20160311, 
#                      LC08_129015_20180317, 
#       
#                                            LC08_128016_20140315, 
#                      LC08_128016_20140331, 
#                      LC08_128016_20160304, 
#                      LC08_128016_20180310, LC08_129016_20140306, 
#                      LC08_129016_20160311, 
#                      LC08_129016_20180301, 
#                      LC08_129016_20180317, 
# Calculate mean
#l5_08_mean <- calc(l5_08, mean, na.rm=T)
#l7_08_mean <- calc(l7_08, mean, na.rm=T)
#l8_08_mean <- calc(l8_08, mean, na.rm=T)
lall_03_mean <- calc(lall_03, mean, na.rm=T)

# Calculate sd
#l5_08_sd <- calc(l5_08, sd, na.rm=T)
#l7_08_sd <- calc(l7_08, sd, na.rm=T)
#l8_08_sd <- calc(l8_08, sd, na.rm=T)
lall_03_sd <- calc(lall_03, sd, na.rm=T)


#l5_08_emsi <- overlay(l5_08, l5_08_mean, l5_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

#l7_08_emsi <- overlay(l7_08, l7_08_mean, l7_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

#l8_08_emsi <- overlay(l8_08, l5_08_mean, l5_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

lall_03_emsi <- overlay(lall_03, lall_03_mean, lall_03_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```

```{r}
#l5_09 <- brick()

# Landsat 8 September dates
#l8_09 <- brick(LC08_035038_20130924,
#                       LC08_035038_20140911,
#                       LC08_035038_20150930,
#                       LC08_035038_20160916,
#                       LC08_035038_20180906)
# Calculate mean
#l8_09_mean <- calc(l8_09, mean)
# Calculate sd
#l8_09_sd <- calc(l8_09, sd)
#l8_09_emsi <- overlay(l8_09, l8_09_mean, l8_09_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```


# Create Leaflet Map of study area 
https://rstudio.github.io/leaflet
http://leafletjs.com/
https://www.r-bloggers.com/interactive-mapping-with-leaflet-in-r/
https://www.color-hex.com/color-palette/19447 

We are going to use a topo map, overlayed with a street map to show states.
To browse all the provider layers,
see http://leaflet-extras.github.io/leaflet-providers/preview/index.html

```{r message=FALSE, warning=FALSE}
# Create custom NDVI color pallete
pal1 <- colorNumeric(c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), values(lall_03_mean), na.color = "transparent")

pal <- colorNumeric(c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), values(lall_03_emsi[[1]]), na.color = "transparent")
```

```{r}
# Okavango Early Period 
m <- leaflet() %>%
  addTiles() %>%
  #addLegend(pal = pal, values = values(lall_03_emsi[[1]]), title = "EMSI") %>%
  #addLegend(pal = pal1, values = values(lall_03_mean), title = "NDVI") %>%
  addRasterImage(lall_03_mean, group = "March Mean NDVI", colors = pal1, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_sd, group = "March Standard Deviation NDVI", colors = pal1, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[1]], group = "March 1987 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[2]], group = "March 1989 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[3]], group = "March 1990 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[4]], group = "March 1991 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[5]], group = "March 1992 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[6]], group = "March 1993 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%  
  addRasterImage(lall_03_emsi[[7]], group = "March 1994 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[8]], group = "March 1995 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%  
  addRasterImage(lall_03_emsi[[9]], group = "March 1996 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[10]], group = "March 1997 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[11]], group = "March 1998 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
setView(lng = -110.55, lat = 31.3, zoom = 12) %>%
addProviderTiles("Stamen.Toner", group = "Stamen") %>%
addProviderTiles("Esri.WorldImagery", group = "ESRI Satellite", options = providerTileOptions(opacity = 0.66, transparent = TRUE)) %>%
addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
#layers control panel
addLayersControl(baseGroups = c("Stamen", "ESRI Satellite", "OpenStreetMap"), overlayGroups = c("March Mean NDVI", "March Standard Deviation NDVI", "March 1987 EMSI", "March 1989 EMSI", "March 1990 EMSI", "March 1991 EMSI", "March 1992 EMSI", "March 1993 EMSI", "March 1994 EMSI", "March 1995 EMSI", "March 1996 EMSI", "March 1997 EMSI", "March 1998 EMSI"), options = layersControlOptions(collapsed = TRUE))

m
```


```{r}
# Okavango recent period
m <- leaflet() %>%
  addTiles() %>%
  #addLegend(pal = pal, values = values(lall_03_emsi[[1]]), title = "EMSI") %>%
  #addLegend(pal = pal1, values = values(lall_03_mean), title = "NDVI") %>%
  addRasterImage(lall_03_mean, group = "March Mean NDVI", colors = pal1, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_sd, group = "March Standard Deviation NDVI", colors = pal1, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[1]], group = "March 2013 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[3]], group = "March 2014 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[5]], group = "March 2015 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_03_emsi[[6]], group = "March 2016 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
setView(lng = 22.30, lat = -19.10, zoom = 8) %>%
addProviderTiles("Stamen.Toner", group = "Stamen") %>%
addProviderTiles("Esri.WorldImagery", group = "ESRI Satellite", options = providerTileOptions(opacity = 0.66, transparent = TRUE)) %>%
addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
#layers control panel
addLayersControl(baseGroups = c("Stamen","ESRI Satellite", "OpenStreetMap"), overlayGroups = c("March Mean NDVI", "March Standard Deviation NDVI", "March 2013 EMSI", "March 2014 EMSI", "March 2015 EMSI", "March 2016 EMSI"), options = layersControlOptions(collapsed = TRUE))

m
```

```{r}
year_id <- c('LT05_175073_19870330'="1987", 
                      'LT05_175073_19890319'="1989", 
                      'LT05_175073_19900322'="1990", 
                      'LT05_175073_19910309'="1991", 
                      'LT05_175073_19920327'="1992", 
                      'LT05_175073_19930314'="1993", 
                      'LT05_175073_19940317'="1994", 
                      'LT05_175073_19950304'="1995", 
                      'LT05_175073_19960322'="1996", 
                      'LT05_175073_19970309'="1997", 
                      'LE07_175073_20010312'="2001", 
                      'LE07_175073_20020331'="2002", 
                      'LE07_175073_20030318'="2003", 
                      'LT05_175073_20040312'="2004", 
                      'LE07_175073_20050307'="2005", 
                      'LE07_175073_20060326'="2006", 
                      'LE07_175073_20070313'="2007",
                      'LT05_175073_20080307'="2008",
                      'LT05_175073_20090326'="2009",
                      'LE07_175073_20100321'="2010", 
                      'LE07_175073_20130313'="2013",
                      'LC08_175073_20140324'="2014", 
                      'LC08_175073_20150311'="2015",
                      'LC08_175073_20160329'="2016", 
                      'LC08_175073_20170316'="2017", 
                      'LC08_175073_20180319'="2018")
```

```{r}
## Multipanel graph Augusts 1987 - 2018
lall_stack <- stack(LT05_175073_19870330, 
                      LT05_175073_19890319,
                      LT05_175073_19900322, 
                      LT05_175073_19910309,
                      LT05_175073_19920327,
                      LT05_175073_19930314,
                      LT05_175073_19940317,
                      LT05_175073_19950304,
                      LT05_175073_19960322,
                      LT05_175073_19970309,
                      LE07_175073_20010312, 
                      LE07_175073_20020331,
                      LE07_175073_20030318,
                      LT05_175073_20040312,
                      LE07_175073_20050307,
                      LE07_175073_20060326, 
                      LE07_175073_20070313,
                      LT05_175073_20080307,
                      LT05_175073_20090326,
                      LE07_175073_20100321,
                      LE07_175073_20130313,
                      LC08_175073_20140324,
                      LC08_175073_20150311,
                      LC08_175073_20160329,
                      LC08_175073_20170316,
                      LC08_175073_20180319)

lall_stack_df <- as.data.frame(lall_stack, xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = lall_stack_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  facet_wrap(~ variable, labeller = as_labeller(year_ids), ncol = 6) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
 
ggsave('~/Documents/ndvi_timeseries_okavango.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l1987_stack_ndvi_df <- as.data.frame(lall_stack[[1]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l1987_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/ndvi_1987_okavango.png', width = 12, height = 8, dpi = 300z, bg = "transparent")
```
```{r}
l1996_stack_ndvi_df <- as.data.frame(lall_stack[[9]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l1996_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/ndvi_1996_okavango.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```
```{r}
l2016_stack_ndvi_df <- as.data.frame(lall_stack[[24]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2016_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/ndvi_2016_okavango.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```



```{r}
# Calculate mean
lall_stack_mean <- calc(lall_stack, mean, na.rm=T)
# Calculate sd
lall_stack_sd <- calc(lall_stack, sd, na.rm=T)
lall_stack_emsi <- overlay(lall_stack, lall_stack_mean, lall_stack_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```

```{r}
# Rename variable layers back to year dates
year_ids <- c(
              'layer.1'="1987",
              'layer.2'="1989",
              'layer.3'="1990",
              'layer.4'="1991",
              'layer.5'="1992",
              'layer.6'="1993",
              'layer.7'="1994",
              'layer.8'="1995",
              'layer.9'="1996",
              'layer.10'="1997",
              'layer.11'="2001",
              'layer.12'="2002",
              'layer.13'="2003",
              'layer.14'="2004",
              'layer.15'="2005",
              'layer.16'="2006",
              'layer.17'="2007",
              'layer.18'="2008",
              'layer.19'="2009",
              'layer.20'="2010",
              'layer.21'="2013",
              'layer.22'="2014",
              'layer.23'="2015",
              'layer.24'="2016",
              'layer.25'="2017",
              'layer.26'="2018"
              )
```

```{r}
lall_stack_emsi_df <- as.data.frame(lall_stack_emsi, xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = lall_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-2.5,2.7), guide = guide_colorbar(title ="EMSI")) +
  facet_wrap(~ variable, labeller = as_labeller(year_ids), ncol = 6) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_timeseries_okavango.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l1987_stack_emsi_df <- as.data.frame(lall_stack_emsi[[1]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l1987_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3,2.7), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_1987_okavango.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l1996_stack_emsi_df <- as.data.frame(lall_stack_emsi[[9]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l1996_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3.2,2.7), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_1996_okavango.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2016_stack_emsi_df <- as.data.frame(lall_stack_emsi[[24]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2016_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3,3.2), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_2016_okavango.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```


```{r}
ggplot(lall_stack_emsi_df) + geom_histogram(aes(value, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-2.5,2.7), guide = guide_colorbar(title ="EMSI")) +
  ylab("Density") + xlab("EMSI") + ggtitle("March") +
  facet_wrap(~variable, labeller = as_labeller(year_ids), ncol = 6) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
```