---
title: "US Mexico EMSI Raster Calculation"
author: "Tyson Lee Swetnam"
date: "10/09/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Linux Dependencies

Instructions for installing geospatial packages are provided by [The Carpentries](https://datacarpentry.org/geospatial-workshop/setup.html)

```{bash}
# sudo add-apt-repository ppa:ubuntugis
# sudo apt-get update
# sudo apt-get install libgdal-dev libgeos-dev libproj-dev
# sudo apt-get install libudunits2-dev
```

# R Libraries

## Install Packages
```{r message=FALSE, warning=FALSE}
# Carpentries recommended raster packages
# update.packages(checkBuilt = TRUE, ask=FALSE)
# install.packages(c('dplyr', 'ggplot2', 'raster', 'rgdal', 'rasterVis', 'remotes', 'sf'))
```

```{r message=FALSE, warning=FALSE}
## Install additional R Packages
# install.packages("maptools")
# install.packages("lattice")
# install.packages("RQGIS")
# install.packages("leaflet")
# install.packages("magrittr")
# install.packages("xlsx")
```
## Load Libraries
```{r message=FALSE, warning=FALSE}
# Load libraries
library(rgdal)
library(grid)
library(lattice)
library(raster) 
library(ggplot2)
library(RColorBrewer)
library(leaflet)
library(magrittr)
library(PerformanceAnalytics)
library(lubridate)
library(stringr)
library(reshape)
library(scales)
library(rasterVis)
library(RColorBrewer)
library(viridis)
```

# Import Rasters from data directory
```{r message=FALSE, warning=FALSE}
# Test for raster metadata with GDALinfo
GDALinfo("~/emsi/data/collections/usmex/gee-collection-usmex-landsat7/LE07_035038_20000912.tif")
```

## Import Raster time series for Landsats 5,7,8
```{r message=FALSE, warning=FALSE}
# Load all rasters in usmex-landsat5
setwd("~/emsi/data/collections/usmex/gee-collection-usmex-landsat5/")
rlist5=list.files(getwd(), pattern="tif$", full.names=FALSE)
for(i in rlist5) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) } 

# Load all rasters in usmex-landsat7
setwd("~/emsi/data/collections/usmex/gee-collection-usmex-landsat7/")
rlist7=list.files(getwd(), pattern="tif$", full.names=FALSE)
for(i in rlist7) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) }

# Load all rasters in usmex-landsat8
setwd("~/emsi/data/collections/usmex/gee-collection-usmex-landsat8/")
rlist8=list.files(getwd(), pattern="tif$", full.names=FALSE)
for(i in rlist8) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) }
```

```{r}
setwd("~/emsi/data/collections/usmex/gee-collection-usmex-landsat5/")
rlist5 <- list.files(getwd(),full.names = TRUE,pattern = ".tif$")
l5_stack <- stack(rlist5)
l5_stack
```

```{r}
list_l5 <- ls(pattern="LT05", all.names = TRUE) 
list_l5
dates_l5 = as.Date(str_sub(list_l5, -8 ,-1), format="%Y%m%d")

list_l7 <- ls(pattern="LE07", all.names = TRUE) 
list_l7
dates_l7 = as.Date(str_sub(list_l7, -8 ,-1), format="%Y%m%d")

list_l8 <- ls(pattern="LC08", all.names = TRUE) 
list_l8
dates_l8 = as.Date(str_sub(list_l8, -8 ,-1), format="%Y%m%d")

list_09 <- ls(pattern = "09", all.names = TRUE)
```

# August EMSI calc prep
```{r}
# Landsat 5,7,8 August scenes
l5_08 <- brick(LT05_035038_19850826,
                        LT05_035038_19870816,
                        LT05_035038_19880818,
                        LT05_035038_19890821,
                        LT05_035038_19900824,
                        LT05_035038_19910811,
                        LT05_035038_19920813,
                        LT05_035038_19930816,
                        LT05_035038_19940803,
                        LT05_035038_19950806,
                        LT05_035038_19950822,
                        LT05_035038_19960824,
                        LT05_035038_19970827,
                        LT05_035038_19980814,
                        LT05_035038_19980830,
                        LT05_035038_19990817,
                        LT05_035038_20000803,
                        LT05_035038_20000819,
                        LT05_035038_20010822,
                        LT05_035038_20020809,
                        LT05_035038_20020825,
                        LT05_035038_20030812,
                        LT05_035038_20040830,
                        LT05_035038_20050801,
                        LT05_035038_20070823,
                        LT05_035038_20090828,
                        LT05_035038_20100815,
                        LT05_035038_20110802,
                        LT05_035038_20110818)
               
l7_08 <- brick(LE07_035038_19990809,
                        LE07_035038_19990825,
                        LE07_035038_20020817,
                        LE07_035038_20030804,
                        LE07_035038_20040806,
                        LE07_035038_20040822,
                        LE07_035038_20050825,
                        LE07_035038_20060828,
                        LE07_035038_20070815,
                        LE07_035038_20070831,
                        LE07_035038_20080801,
                        LE07_035038_20090804,
                        LE07_035038_20100823,
                        LE07_035038_20110826,
                        LE07_035038_20120812,
                        LE07_035038_20160807,
                        LE07_035038_20180829)
            
l8_08 <- brick( LC08_035038_20150813,
                        LC08_035038_20150829,
                        LC08_035038_20160815,
                        LC08_035038_20180805)

lall_08 <- brick(LT05_035038_19850826,
                        LT05_035038_19870816,
                        LT05_035038_19880818,
                        LT05_035038_19890821,
                        LT05_035038_19900824,
                        LT05_035038_19910811,
                        LT05_035038_19920813,
                        LT05_035038_19930816,
                        LT05_035038_19940803,
                        LT05_035038_19950806,
                        LT05_035038_19950822,
                        LT05_035038_19960824,
                        LT05_035038_19970827,
                        LT05_035038_19980814,
                        LT05_035038_19980830,
                        LT05_035038_19990817,
                        LT05_035038_20000803,
                        LT05_035038_20000819,
                        LT05_035038_20010822,
                        LT05_035038_20020809,
                        LT05_035038_20020825,
                        LT05_035038_20030812,
                        LT05_035038_20040830,
                        LT05_035038_20050801,
                        LT05_035038_20070823,
                        LT05_035038_20090828,
                        LT05_035038_20100815,
                        LT05_035038_20110802,
                        LT05_035038_20110818,
                        LE07_035038_19990809,
                        LE07_035038_19990825,
                        LE07_035038_20020817,
                        LE07_035038_20030804,
                        LE07_035038_20040806,
                        LE07_035038_20040822,
                        LE07_035038_20050825,
                        LE07_035038_20060828,
                        LE07_035038_20070815,
                        LE07_035038_20070831,
                        LE07_035038_20080801,
                        LE07_035038_20090804,
                        LE07_035038_20100823,
                        LE07_035038_20110826,
                        LE07_035038_20120812,
                        LE07_035038_20160807,
                        LE07_035038_20180829,
                        LC08_035038_20150813,
                        LC08_035038_20150829,
                        LC08_035038_20160815,
                        LC08_035038_20180805)

# Calculate mean
l5_08_mean <- calc(l5_08, mean, na.rm=T)
l7_08_mean <- calc(l7_08, mean, na.rm=T)
l8_08_mean <- calc(l8_08, mean, na.rm=T)
lall_08_mean <- calc(lall_08, mean, na.rm=T)

# Calculate sd
l5_08_sd <- calc(l5_08, sd, na.rm=T)
l7_08_sd <- calc(l7_08, sd, na.rm=T)
l8_08_sd <- calc(l8_08, sd, na.rm=T)
lall_08_sd <- calc(lall_08, sd, na.rm=T)


l5_08_emsi <- overlay(l5_08, l5_08_mean, l5_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

l7_08_emsi <- overlay(l7_08, l7_08_mean, l7_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

l8_08_emsi <- overlay(l8_08, l5_08_mean, l5_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

lall_08_emsi <- overlay(lall_08, lall_08_mean, lall_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```

```{r}
l5_09 <- brick()

# Landsat 8 September dates
l8_09 <- brick(LC08_035038_20130924,
                       LC08_035038_20140911,
                       LC08_035038_20150930,
                       LC08_035038_20160916,
                       LC08_035038_20180906)
# Calculate mean
l8_09_mean <- calc(l8_09, mean)
# Calculate sd
l8_09_sd <- calc(l8_09, sd)
l8_09_emsi <- overlay(l8_09, l8_09_mean, l8_09_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```


# Create Leaflet Map of study area 
https://rstudio.github.io/leaflet
http://leafletjs.com/
https://www.r-bloggers.com/interactive-mapping-with-leaflet-in-r/
https://www.color-hex.com/color-palette/19447 

We are going to use a topo map, overlayed with a street map to show states.
To browse all the provider layers,
see http://leaflet-extras.github.io/leaflet-providers/preview/index.html


```{r}
m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852,
    popup="The birthplace of R")
m  # Print the map
```

```{r message=FALSE, warning=FALSE}
# Create custom NDVI color pallete
pal1 <- colorNumeric(c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), values(lall_08_mean), na.color = "transparent")

pal <- colorNumeric(c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), values(lall_08_emsi[[1]]), na.color = "transparent")
```

```{r}
m <- leaflet() %>%
  addTiles() %>%
  #addLegend(pal = pal, values = values(l8_08_emsi[[1]]), title = "EMSI") %>%
  #addLegend(pal = pal1, values = values(l8_08_mean), title = "NDVI") %>%
  addRasterImage(l8_08_mean, group = "August Mean NDVI", colors = pal1, opacity = 1.0) %>%
  addRasterImage(l8_08_sd, group = "August Standard Deviation NDVI", colors = pal1, opacity = 1.0) %>%
  addRasterImage(l8_08_emsi[[1]], group = "early August 2015 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(l8_08_emsi[[2]], group = "August 2015 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(l8_08_emsi[[3]], group = "August 2016 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(l8_08_emsi[[4]], group = "August 2018 EMSI", colors = pal, opacity = 1.0) %>%
setView(lng = -110.55, lat = 31.3, zoom = 12) %>%
addProviderTiles("Stamen.Toner", group = "Stamen") %>%
addProviderTiles("Esri.WorldImagery", group = "ESRI Satellite", options = providerTileOptions(opacity = 0.66, transparent = TRUE)) %>%
addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
#layers control panel
addLayersControl(baseGroups = c("Stamen", "ESRI Satellite", "OpenStreetMap"), overlayGroups = c("August Mean NDVI", "August Standard Deviation NDVI", "early August 2015 EMSI", "August 2015 EMSI", "August 2016 EMSI", "August 2018 EMSI"), options = layersControlOptions(collapsed = TRUE))

m
```

```{r}
# Landsat 5
m <- leaflet() %>%
  addTiles() %>%
  addLegend(pal = pal, values = values(l5_08_emsi[[1]]), title = "EMSI") %>%
  addLegend(pal = pal1, values = values(l5_08_mean), title = "NDVI") %>%
  addRasterImage(l5_08_mean, group = "August Mean NDVI", colors = pal1, opacity = 1.0) %>%
  addRasterImage(l5_08_sd, group = "August Standard Deviation NDVI", colors = pal1, opacity = 1.0) %>%
  addRasterImage(l5_08_emsi[[1]], group = "August 1985 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(l5_08_emsi[[2]], group = "August 1987 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(l5_08_emsi[[3]], group = "August 1988 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(l5_08_emsi[[4]], group = "August 1989 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(l5_08_emsi[[5]], group = "August 1990 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(l5_08_emsi[[6]], group = "August 1991 EMSI", colors = pal, opacity = 1.0) %>%  
  addRasterImage(l5_08_emsi[[7]], group = "August 1992 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(l5_08_emsi[[8]], group = "August 1993 EMSI", colors = pal, opacity = 1.0) %>%  
  addRasterImage(l5_08_emsi[[9]], group = "August 1994 EMSI", colors = pal, opacity = 1.0) %>%
setView(lng = -110.55, lat = 31.3, zoom = 12) %>%
addProviderTiles("Stamen.Toner", group = "Stamen") %>%
addProviderTiles("Esri.WorldImagery", group = "ESRI Satellite", options = providerTileOptions(opacity = 0.66, transparent = TRUE)) %>%
addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
#layers control panel
addLayersControl(baseGroups = c("Stamen", "ESRI Satellite", "OpenStreetMap"), overlayGroups = c("August Mean NDVI", "August Standard Deviation NDVI", "August 1985 EMSI", "August 1987 EMSI", "August 1988 EMSI", "August 1989 EMSI", "August 1990 EMSI", "August 1991 EMSI", "August 1992 EMSI", "August 1993 EMSI", "August 1994 EMSI"), options = layersControlOptions(collapsed = TRUE))

m
```

```{r}

library(raster)
library(rasterVis)
```

```{r}
m <- leaflet() %>%
  addTiles() %>%
  #addLegend(pal = pal, values = values(l8_08_emsi[[1]]), title = "EMSI") %>%
  #addLegend(pal = pal1, values = values(l8_08_mean), title = "NDVI") %>%
  addRasterImage(lall_08_mean, group = "August Mean NDVI", colors = pal1, opacity = 1.0) %>%
  addRasterImage(lall_08_sd, group = "August Standard Deviation NDVI", colors = pal1, opacity = 1.0) %>%
  addRasterImage(lall_08_emsi[[1]], group = "early August 2015 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(lall_08_emsi[[2]], group = "August 2015 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(lall_08_emsi[[3]], group = "August 2016 EMSI", colors = pal, opacity = 1.0) %>%
  addRasterImage(lall_08_emsi[[4]], group = "August 2018 EMSI", colors = pal, opacity = 1.0) %>%
setView(lng = -110.55, lat = 31.3, zoom = 12) %>%
addProviderTiles("Stamen.Toner", group = "Stamen") %>%
addProviderTiles("Esri.WorldImagery", group = "ESRI Satellite", options = providerTileOptions(opacity = 0.66, transparent = TRUE)) %>%
addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
#layers control panel
addLayersControl(baseGroups = c("Stamen", "ESRI Satellite", "OpenStreetMap"), overlayGroups = c("August Mean NDVI", "August Standard Deviation NDVI", "early August 2015 EMSI", "August 2015 EMSI", "August 2016 EMSI", "August 2018 EMSI"), options = layersControlOptions(collapsed = TRUE))

m
```

```{r}
year_id <- c('LT05_035038_19850826'="1985",
                        'LT05_035038_19870816'="1987",
                        'LT05_035038_19880818'="1985",
                        'LT05_035038_19890821'="1989",
                        'LT05_035038_19900824'="1990",
                        'LT05_035038_19910811'="1991",
                        'LT05_035038_19920813'="1992",
                        'LT05_035038_19930816'="1993",
                        'LT05_035038_19940803'="1994",
                        'LT05_035038_19950806'="1995",
                        'LT05_035038_19960824'="1996",
                        'LT05_035038_19970827'="1997",
                        'LT05_035038_19980814'="1998",
                        'LT05_035038_19990817'="1999",
                        'LT05_035038_20000819'="2000",
                        'LT05_035038_20010822'="2001",
                        'LE07_035038_20020817'="2002",
                        'LT05_035038_20030812'="2003",
                        'LT05_035038_20040830'="2004",
                        'LT05_035038_20050801'="2005",
                        'LE07_035038_20060828'="2006",
                        'LT05_035038_20070823'="2007",
                        'LE07_035038_20080801'="2008",
                        'LT05_035038_20090828'="2009",
                        'LT05_035038_20100815'="2010",
                        'LT05_035038_20110818'="2011",
                        'LE07_035038_20120812'="2012",
                        'LC08_035038_20150813'="2015",
                        'LC08_035038_20160815'="2016",
                        'LC08_035038_20180805'="2018"
                        )
```

```{r}
## Multipanel graph Augusts 1985 - 2018
lall_stack <- stack(LT05_035038_19850826,
                        LT05_035038_19870816,
                        LT05_035038_19880818,
                        LT05_035038_19890821,
                        LT05_035038_19900824,
                        LT05_035038_19910811,
                        LT05_035038_19920813,
                        LT05_035038_19930816,
                        LT05_035038_19940803,
                        LT05_035038_19950806,
                        LT05_035038_19960824,
                        LT05_035038_19970827,
                        LT05_035038_19980814,
                        LT05_035038_19990817,
                        LT05_035038_20000819,
                        LT05_035038_20010822,
                        LE07_035038_20020817,
                        LT05_035038_20030812,
                        LT05_035038_20040830,
                        LT05_035038_20050801,
                        LE07_035038_20060828,
                        LT05_035038_20070823,
                        LE07_035038_20080801,
                        LT05_035038_20090828,
                        LT05_035038_20100815,
                        LT05_035038_20110818,
                        LE07_035038_20120812,
                        LC08_035038_20150813,
                        LC08_035038_20160815,
                        LC08_035038_20180805)

lall_stack_df <- as.data.frame(lall_stack, xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = lall_stack_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  facet_wrap(~ variable, labeller = as_labeller(year_id), ncol = 6) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
 
ggsave('~/Documents/ndvi_timeseries_grassmx.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l1994_stack_ndvi_df <- as.data.frame(lall_stack[[9]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l1994_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/ndvi_1994_grassmx.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```
```{r}
l2001_stack_ndvi_df <- as.data.frame(lall_stack[[16]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2001_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/ndvi_2001_grassmx.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```
```{r}
l2016_stack_ndvi_df <- as.data.frame(lall_stack[[29]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2016_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/ndvi_2016_grassmx.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```



```{r}
# Calculate mean
lall_stack_mean <- calc(lall_stack, mean, na.rm=T)
# Calculate sd
lall_stack_sd <- calc(lall_stack, sd, na.rm=T)
lall_stack_emsi <- overlay(lall_stack, lall_stack_mean, lall_stack_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```

```{r}
# Rename variable layers back to year dates
year_ids <- c(
              'layer.1'="1985",
              'layer.2'="1987",
              'layer.3'="1988",
              'layer.4'="1989",
              'layer.5'="1990",
              'layer.6'="1991",
              'layer.7'="1992",
              'layer.8'="1993",
              'layer.9'="1994",
              'layer.10'="1995",
              'layer.11'="1996",
              'layer.12'="1997",
              'layer.13'="1998",
              'layer.14'="1999",
              'layer.15'="2000",
              'layer.16'="2001",
              'layer.17'="2002",
              'layer.18'="2003",
              'layer.19'="2004",
              'layer.20'="2005",
              'layer.21'="2006",
              'layer.22'="2007",
              'layer.23'="2008",
              'layer.24'="2009",
              'layer.25'="2010",
              'layer.26'="2011",
              'layer.27'="2012",
              'layer.28'="2015",
              'layer.29'="2016",
              'layer.30'="2018"
              )
```

```{r}
lall_stack_emsi_df <- as.data.frame(lall_stack_emsi, xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = lall_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-2.5,2.7), guide = guide_colorbar(title ="EMSI")) +
  facet_wrap(~ variable, labeller = as_labeller(year_ids), ncol = 6) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_timeseries_grassmx.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l1994_stack_emsi_df <- as.data.frame(lall_stack_emsi[[9]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l1994_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3,2.7), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_1994_grassmx.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2001_stack_emsi_df <- as.data.frame(lall_stack_emsi[[16]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2001_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3,2.7), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_2001_grassmx.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```
```{r}
l2016_stack_emsi_df <- as.data.frame(lall_stack_emsi[[29]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2016_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3,2.7), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_2016_grassmx.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```


```{r}
ggplot(lall_stack_emsi_df) + geom_histogram(aes(value, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-2.5,2.7), guide = guide_colorbar(title ="EMSI")) +
  ylab("Density") + xlab("EMSI") + ggtitle("August") +
  facet_wrap(~variable, labeller = as_labeller(year_ids), ncol = 6) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
```