---
title: "Yakutia EMSI Raster Calculation"
author: "Tyson Lee Swetnam"
date: "2021-03-03"
output: html_document
---

# Setup

## Linux Dependencies

If you are running this for the first time, you may need to install additional geospatial dependencies

```{bash message=FALSE, warning=FALSE}
# sudo add-apt-repository ppa:ubuntugis
# sudo apt-get update
# sudo apt-get install libgdal-dev libgeos-dev libproj-dev
# sudo apt-get install libudunits2-dev
```

## Install missing R Libraries

The following code block executes when the notebook is opened. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("dplyr")) {
   install.packages("dplyr")
   library(dplyr)
}
if (!require("ggplot2")) {
   install.packages("ggplot2")
   library(ggplot2)
}  
if (!require("grid")) {
   install.packages("grid")
   library(grid)
} 
if (!require("lattice")) {
   install.packages("lattice")
   library(lattice)
}
if (!require("leaflet")) {
   install.packages("leaflet")
   library(leaflet)
}
if (!require("lubridate")) {
   install.packages("lubridate")
   library(lubridate)
}
if (!require("magrittr")) {
   install.packages("magrittr")
   library(magrittr)
}
if (!require("raster")) {
   install.packages("raster")
   library(raster)  
}
if (!require("rasterVis")) {
   install.packages("rasterVis")
   library(rasterVis)   
}
if (!require("RColorBrewer")) {
   install.packages("RColorBrewer")
   library(RColorBrewer)
}
if (!require("reshape")) {
   install.packages("reshape")
   library(reshape)
}
if (!require("rgdal")) {
   install.packages("rgdal")
   library(rgdal)   
}  
if (!require("remotes")) {
   install.packages("remotes")
   library(remotes)   
}  
if (!require("PerformanceAnalytics")) {
   install.packages("PerformanceAnalytics")
   library(PerformanceAnalytics)
}
if (!require("scales")) {
   install.packages("scales")
   library(scales)
}
if (!require("sf")) {
   install.packages("sf")
   library(sf)   
}
if (!require("stringr")) {
   install.packages("stringr")
   library(stringr)
}
if (!require("viridis")) {
   install.packages("viridis")
   library(viridis)
}
```

# Transfer datasets from CyVerse

This command uses `wget` to download data into the container.

The CyVerse iRODS data store also supports [WebDav](https://data.cyverse.org); we are hosting the extracted raster data in a folder with anonymous public read-only settings.

All of the landsat data are ~22.8 GB in size, and will take ~18 minutes to transfer within the CyVerse workbench for all four sites.

A subset of the collection will take less time, here we're taking the `/okavango` collection. 

```{bash message=FALSE, warning=FALSE}
# uncomment to run
 cd ~/emsi/data/
time wget -r -nH --cut-dirs=7 --no-parent -l8 --reject="index.html*" https://data.cyverse.org/dav-anon/iplant/home/tswetnam/emsi/data/collections/okavango
```

# Import Raster from data directory and check the header
```{r message=FALSE, warning=FALSE}
# Test for raster metadata with GDALinfo
GDALinfo("~/emsi/data/collections/yakutia/gee-collection-yakutia-landsat8/LC08_175073_20171026.tif")
```

## Import Raster time series for Landsats 5,7,8
```{r message=FALSE, warning=FALSE}
# Load all rasters in yakutia-landsat5
setwd("~/emsi/data/collections/yakutia/gee-collection-yakutia-landsat5/")
rlist5=list.files(getwd(), pattern=".tif$", full.names=FALSE)
for(i in rlist5) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) } 

# Load all rasters in yakutia-landsat7
setwd("~/emsi/data/collections/yakutia/gee-collection-yakutia-landsat7/")
rlist7=list.files(getwd(), pattern="tif$", full.names=FALSE)
for(i in rlist7) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) }

# Load all rasters in yakutia-landsat8
setwd("~/emsi/data/collections/yakutia/gee-collection-yakutia-landsat8/")
rlist8=list.files(getwd(), pattern="tif$", full.names=FALSE)
for(i in rlist8) { assign(unlist(strsplit(i, "[.]"))[1], raster(i)) }
```

```{r}
list_l5 <- ls(pattern="LT05", all.names = TRUE) 
dates_l5 = as.Date(str_sub(list_l5, -8 ,-1), format="%Y%m%d")

list_l7 <- ls(pattern="LE07", all.names = TRUE) 
dates_l7 = as.Date(str_sub(list_l7, -8 ,-1), format="%Y%m%d")

list_l8 <- ls(pattern="LC08", all.names = TRUE) 
dates_l8 = as.Date(str_sub(list_l8, -8 ,-1), format="%Y%m%d")

list_06 <- ls(pattern = "06", all.names = TRUE)
list_06

list_07 <- ls(pattern = "07", all.names = TRUE)
list_07
```

# June (peak greenness) EMSI calc prep
```{r}
      
lall_06 <- brick(LT05_128016_19990728,
                  LT05_128016_20010701,
                  LT05_128016_20010717,
                  LT05_128016_20020704,
                  LT05_128016_20020720,
                  LE07_128016_20030715,
                  LT05_128016_20040725,
                  LE07_128016_20050704,
                  LE07_128016_20060520,
                  LE07_128016_20060723, 
                  LT05_128016_20060731,
                  LT05_128016_20070515,
                  LE07_128016_20070726,
                  LE07_128016_20080626,
                  LE07_128016_20080813,
                  LE07_128016_20090715,
                  LE07_128016_20090917,
                  LE07_128016_20100819,
                  LE07_128016_20110603,
                  LT05_128016_20110713,
                  LE07_128016_20120707,
                  LE07_128016_20120723,
                  LE07_128016_20130726,
                  LE07_128016_20140611,
                  LC08_128016_20140619,
                  LC08_128016_20150606,
                  LC08_128016_20150622,
                  LE07_128016_20160803,
                  LE07_128016_20160920,
                  LC08_128016_20160608,
                  LC08_128016_20160624,
                  LC08_128016_20170611,
                  LC08_128016_20170627,
                  LE07_128016_20170705,
                  LE07_128016_20170721,
                  LC08_128016_20180614,
                  LC08_128016_20180630,
                  LE07_128016_20180521,
                  LE07_128016_20180606
                  )
    

# Landsat 5,7,8 June
lall_06 <- brick(LT05_127016_20000605, 
                 LT05_127016_20010624,
                 LT05_127016_20020627,
                 LE07_127016_20050611,
                 LT05_127016_20060910,
                 LT05_127016_20070812,
                 LT05_127016_20090630, 
                 LT05_127016_20100601,
                 LE07_127016_20120630, 
                 LC08_127016_20130609)

                 
# Calculate mean
#l5_08_mean <- calc(l5_08, mean, na.rm=T)
#l7_08_mean <- calc(l7_08, mean, na.rm=T)
#l8_08_mean <- calc(l8_08, mean, na.rm=T)
lall_06_mean <- calc(lall_06, mean, na.rm=T)

# Calculate sd
#l5_08_sd <- calc(l5_08, sd, na.rm=T)
#l7_08_sd <- calc(l7_08, sd, na.rm=T)
#l8_08_sd <- calc(l8_08, sd, na.rm=T)
lall_06_sd <- calc(lall_06, sd, na.rm=T)


#l5_08_emsi <- overlay(l5_08, l5_08_mean, l5_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

#l7_08_emsi <- overlay(l7_08, l7_08_mean, l7_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

#l8_08_emsi <- overlay(l8_08, l5_08_mean, l5_08_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })

lall_06_emsi <- overlay(lall_06, lall_06_mean, lall_06_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```

```{r}
#l5_09 <- brick()

# Landsat 8 September dates
#l8_09 <- brick(LC08_035038_20130924,
#                       LC08_035038_20140911,
#                       LC08_035038_20150930,
#                       LC08_035038_20160916,
#                       LC08_035038_20180906)
# Calculate mean
#l8_09_mean <- calc(l8_09, mean)
# Calculate sd
#l8_09_sd <- calc(l8_09, sd)
#l8_09_emsi <- overlay(l8_09, l8_09_mean, l8_09_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```


# Create Leaflet Map of study area 
https://rstudio.github.io/leaflet
http://leafletjs.com/
https://www.r-bloggers.com/interactive-mapping-with-leaflet-in-r/
https://www.color-hex.com/color-palette/19447 

We are going to use a topo map, overlayed with a street map to show states.
To browse all the provider layers,
see http://leaflet-extras.github.io/leaflet-providers/preview/index.html

```{r message=FALSE, warning=FALSE}
# Create custom NDVI color pallete
pal1 <- colorNumeric(c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), values(lall_06_mean), na.color = "transparent")

pal <- colorNumeric(c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), values(lall_06_emsi[[1]]), na.color = "transparent")
```

```{r}
# Scene 127016
m <- leaflet() %>%
  addTiles() %>%
  #addLegend(pal = pal, values = values(lall_06_emsi[[1]]), title = "EMSI") %>%
  #addLegend(pal = pal1, values = values(lall_06_mean), title = "NDVI") %>%
  addRasterImage(lall_06_mean, group = "June Mean NDVI", colors = pal1, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_sd, group = "June Standard Deviation NDVI", colors = pal1, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[1]], group = "June 2000 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[2]], group = "June 2001 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[3]], group = "June 2002 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[4]], group = "June 2005 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[5]], group = "July 2006 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[6]], group = "July 2007 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%  
  addRasterImage(lall_06_emsi[[7]], group = "July 2009 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[8]], group = "July 2010 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%  
  addRasterImage(lall_06_emsi[[9]], group = "July 2012 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[10]], group = "July 2013 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
setView(lng = 120.3300, lat =  63.4300, zoom = 12) %>%
addProviderTiles("Stamen.Toner", group = "Stamen") %>%
addProviderTiles("Esri.WorldImagery", group = "ESRI Satellite", options = providerTileOptions(opacity = 0.66, transparent = TRUE)) %>%
addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
#layers control panel
addLayersControl(baseGroups = c("Stamen", "ESRI Satellite", "OpenStreetMap"), overlayGroups = c("June Mean NDVI", "June Standard Deviation NDVI", "June 2000 EMSI", "June 2001 EMSI", "June 2002 EMSI", "June 2005 EMSI", "June 2006 EMSI", "July 2007 EMSI", "July 2009 EMSI", "July 2010 EMSI", "July 2012 EMSI", "July 2013 EMSI"), options = layersControlOptions(collapsed = TRUE))

m
```

```{r}
# Scene 127016
m <- leaflet() %>%
  addTiles() %>%
  #addLegend(pal = pal, values = values(lall_06_emsi[[1]]), title = "EMSI") %>%
  #addLegend(pal = pal1, values = values(lall_06_mean), title = "NDVI") %>%
  addRasterImage(lall_06_mean, group = "June Mean NDVI", colors = pal1, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_sd, group = "June Standard Deviation NDVI", colors = pal1, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[21]], group = "June 2013 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[22]], group = "June 2014a EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[23]], group = "June 2014b EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[24]], group = "June 2015 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[25]], group = "July 2015a EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[26]], group = "July 2015b EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%  
  addRasterImage(lall_06_emsi[[27]], group = "July 2016 EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
  addRasterImage(lall_06_emsi[[28]], group = "July 2017a EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%  
  addRasterImage(lall_06_emsi[[29]], group = "July 2017b EMSI", colors = pal, opacity = 1.0, maxBytes = 32 * 1024 * 1024) %>%
setView(lng = 120.3300, lat =  63.4300, zoom = 12) %>%
addProviderTiles("Stamen.Toner", group = "Stamen") %>%
addProviderTiles("Esri.WorldImagery", group = "ESRI Satellite", options = providerTileOptions(opacity = 0.66, transparent = TRUE)) %>%
addProviderTiles("OpenStreetMap.Mapnik", group = "OpenStreetMap") %>%
#layers control panel
addLayersControl(baseGroups = c("Stamen", "ESRI Satellite", "OpenStreetMap"), overlayGroups = c("June Mean NDVI", "June Standard Deviation NDVI", "June 2013 EMSI", "June 2014a EMSI", "June 2014b EMSI", "June 2015 EMSI", "July 2015a EMSI", "July 2015b EMSI", "July 2016 EMSI", "July 2017a EMSI", "July 2017b EMSI"), options = layersControlOptions(collapsed = TRUE))

m
```


```{r}

year_id <- c('LT05_128016_19990728' = "1999",
                  'LT05_128016_20010701' = "2001",
                  'LT05_128016_20010717' = "2001",
                  'LT05_128016_20020704' = "2002",
                  'LT05_128016_20020720' = "2002",
                  'LE07_128016_20030715' = "2003",
                  'LT05_128016_20040725' = "2004",
                  'LE07_128016_20050704' = "2005",
                  'LE07_128016_20060520' = "2006",
                  'LE07_128016_20060723' = "2006", 
                  'LT05_128016_20060731' = "2006",
                  'LT05_128016_20070515' = "2007",
                  'LE07_128016_20070726' = "2007",
                  'LE07_128016_20080626' = "2008",
                  'LE07_128016_20080813' = "2008",
                  'LE07_128016_20090715' = "2009",
                  'LE07_128016_20090917' = "2009",
                  'LE07_128016_20100819' = "2010",
                  'LE07_128016_20110603' = "2010",
                  'LT05_128016_20110713' = "2011",
                  'LE07_128016_20120707' = "2012",
                  'LE07_128016_20120723' = "2012",
                  'LE07_128016_20130726' = "2013",
                  'LE07_128016_20140611' = "2014",
                  'LC08_128016_20140619' = "2014",
                  'LC08_128016_20150606' = "2015",
                  'LC08_128016_20150622' = "2015",
                  'LE07_128016_20160803' = "2016",
                  'LE07_128016_20160920' = "2016",
                  'LC08_128016_20160608' = "2016",
                  'LC08_128016_20160624' = "2016",
                  'LC08_128016_20170611' = "2017",
                  'LC08_128016_20170627' = "2017",
                  'LE07_128016_20170705' = "2017",
                  'LE07_128016_20170721' = "2017",
                  'LC08_128016_20180614' = "2018",
                  'LC08_128016_20180630' = "2018",
                  'LE07_128016_20180521' = "2018",
                  'LE07_128016_2018060' = "2018")

year_ids <- c('layer.1'="1999",
              'layer.2'="2001",
              'layer.3'="2001",
              'layer.4'="2002",
              'layer.5'="2002",
              'layer.6'="2003",
              'layer.7'="2004",
              'layer.8'="2005",
              'layer.9'="2006",
              'layer.10'="2006",
              'layer.11'="2006",
              'layer.12'="2007",
              'layer.13'="2007",
              'layer.14'="2008",
              'layer.15'="2008",
              'layer.16'="2009",
              'layer.17'="2009",
              'layer.18'="2010",
              'layer.19'="2011",
              'layer.20'="2011",
              'layer.21'="2012",
              'layer.22'="2012",
              'layer.23'="2013",
              'layer.24'="2014",
              'layer.25'="2014",
              'layer.26'="2015",
              'layer.27'="2016",
              'layer.28'="2016",
              'layer.29'="2016",
              'layer.30'="2016",
              'layer.31'="2017",
              'layer.32'="2017",
              'layer.33'="2017",
              'layer.34'="2017",
              'layer.35'="2018",
              'layer.36'="2018",
              'layer.37'="2018",
              'layer.38'="2018")
```

```{r}
## Multipanel graph Summer 2000 - 2013
lall_stack <- stack(LT05_127016_20000605, 
                 LT05_127016_20010624,
                 LT05_127016_20020627,
                 LE07_127016_20050611,
                 LT05_127016_20060910,
                 LT05_127016_20070812,
                 LT05_127016_20090630, 
                 LT05_127016_20100601,
                 LE07_127016_20120630, 
                 LC08_127016_20130609)

lall_stack_df <- as.data.frame(lall_stack, xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = lall_stack_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  facet_wrap(~ variable, labeller = as_labeller(year_id), ncol = 6) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
 
ggsave('~/Documents/ndvi_timeseries_yakutia.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
## Multipanel graph Augusts 2000 - 2018
lall_stack <- stack(LT05_128016_19990728,
                  LT05_128016_20010701,
                  LT05_128016_20010717,
                  LT05_128016_20020704,
                  LT05_128016_20020720,
                  LE07_128016_20030715,
                  LT05_128016_20040725,
                  LE07_128016_20050704,
                  LE07_128016_20060520,
                  LE07_128016_20060723, 
                  LT05_128016_20060731,
                  LT05_128016_20070515,
                  LE07_128016_20070726,
                  LE07_128016_20080626,
                  LE07_128016_20080813,
                  LE07_128016_20090715,
                  LE07_128016_20090917,
                  LE07_128016_20100819,
                  LE07_128016_20110603,
                  LT05_128016_20110713,
                  LE07_128016_20120707,
                  LE07_128016_20120723,
                  LE07_128016_20130726,
                  LE07_128016_20140611,
                  LC08_128016_20140619,
                  LC08_128016_20150606,
                  LC08_128016_20150622,
                  LE07_128016_20160803,
                  LE07_128016_20160920,
                  LC08_128016_20160608,
                  LC08_128016_20160624,
                  LC08_128016_20170611,
                  LC08_128016_20170627,
                  LE07_128016_20170705,
                  LE07_128016_20170721,
                  LC08_128016_20180614,
                  LC08_128016_20180630,
                  LE07_128016_20180521,
                  LE07_128016_20180606)

lall_stack_df <- as.data.frame(lall_stack, xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = lall_stack_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  facet_wrap(~ variable, labeller = as_labeller(year_id), ncol = 6) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
 
ggsave('~/Documents/ndvi_timeseries_yakutia.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2000_stack_ndvi_df <- as.data.frame(lall_stack[[1]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2000_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/ndvi_2000_yakutia.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l1996_stack_ndvi_df <- as.data.frame(lall_stack[[9]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l1996_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/ndvi_1996_yakutia.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2016_stack_ndvi_df <- as.data.frame(lall_stack[[24]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2016_stack_ndvi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(0,0.95), guide = guide_colorbar(title ="NDVI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/ndvi_2016_yakutia.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```



```{r}
# Calculate mean
lall_stack_mean <- calc(lall_stack, mean, na.rm=T)
# Calculate sd
lall_stack_sd <- calc(lall_stack, sd, na.rm=T)
lall_stack_emsi <- overlay(lall_stack, lall_stack_mean, lall_stack_sd, fun = function(r1, r2, r3) { return( (r1 - r2)/r3) })
```

```{r}
# Rename variable layers back to year dates
year_ids <- c('layer.1'="2001",
              'layer.2'="2002",
              'layer.3'="2005",
              'layer.4'="2006",
              'layer.5'="2009",
              'layer.6'="2010",
              'layer.7'="2012",
              'layer.8'="2013")
```

```{r}
lall_stack_emsi_df <- as.data.frame(lall_stack_emsi, xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = lall_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-2.5,2.7), guide = guide_colorbar(title ="EMSI")) +
  facet_wrap(~ variable, labeller = as_labeller(year_ids), ncol = 6) + 
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_timeseries_yakutia.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2010_stack_emsi_df <- as.data.frame(lall_stack_emsi[[8]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2010_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3,2.7), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_2010_yakutia.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2007_stack_emsi_df <- as.data.frame(lall_stack_emsi[[6]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2007_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3.2,2.7), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_2007_yakutia.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```

```{r}
l2016_stack_emsi_df <- as.data.frame(lall_stack_emsi[[24]], xy = TRUE) %>%
    melt(id.vars = c('x','y'))

ggplot() +
  geom_raster(data = l2016_stack_emsi_df, aes(x = x, y = y, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-3,3.2), guide = guide_colorbar(title ="EMSI")) +
  theme(axis.title = element_blank(),
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

ggsave('~/Documents/emsi_2016_yakutia.png', width = 12, height = 8, dpi = 300, bg = "transparent")
```


```{r}
ggplot(lall_stack_emsi_df) + geom_histogram(aes(value, fill = value)) +
  scale_fill_gradientn(colours=c("#6E462C", "#9C8448", "#CCCC66", "#9CAB68", "#306466"), limits=c(-2.5,2.7), guide = guide_colorbar(title ="EMSI")) +
  ylab("Density") + xlab("EMSI") + ggtitle("June") +
  facet_wrap(~variable, labeller = as_labeller(year_ids), ncol = 6) +
  theme_bw() +
  theme(axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())
```