---
title: "Global Map - Figure 3"
author: "Tyson Lee Swetnam"
date: "12/5/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dependencies

This Rmd was created in RStudio-Server, R v3.6.0. It should be usable via the Rocker Geospatial RStudio (https://www.rocker-project.org/images/, http://doi.org/10.5281/zenodo.3384215) Docker container.

Additional dependencies required for generating this figure, uncomment `#` and Run Current Chunk to install:

```{r}
# Install Packages
#=================
install.packages(c("ggplot2", "ggmap", "devtools", "dplyr", "stringr", "ggrepel"))
#install.packages(c("maps", "mapdata"))
#=================
```

```{r}
library(ggplot2)
library(ggmap)
library(maps)
library(mapdata)
library(mapproj)
library(ggrepel)
```

## Global Map

```{r}
#==============
# GET WORLD MAP
#==============

map.world <- map_data("world")
```

```{r}
# Create study area locations
study_areas=data.frame("location" = c("Okavango Delta, Botswana", "Yakutia Taiga, Russian Federation", "USA-Mexico Borderlands", "Acre Rainforest-Agriculture, Brazil"), "long" = c(22.5601, 120.3300, -110.5901, -71.9610), "lat" = c(-19.0901, 63.4300, 31.3400, -9.1901))
```

```{r}
# Generate map and save it

ggplot() + geom_polygon(data = map.world,  aes(x=long, y = lat, group = group), fill = "gray", color = "black", size = 0.15) + 
  geom_point(data = study_areas, aes(x=long, y = lat, label=location), color = "black", size=2) +
  coord_quickmap() +
  coord_fixed(1) +
  xlab("Longitude") +
  ylab("Latitude") +
  geom_label_repel(data = study_areas, aes(x=long, y=lat, label = location),
                  box.padding   = 0, 
                  point.padding = 1,
                  segment.color = 'grey50') +
  theme_minimal()

ggsave('~/emsi/rmd/figures/world_map_fig3.png', width = 8.5, dpi = 300, bg = "transparent")
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
